### 2017-02-14 r_scripts/oregon_clean.rmd
Create oregon_clean.rmd

---
title: "Oregon_Wildfire_Claim"
author: "Jingyang Liu"
date: "February 14, 2017"
output: html_document
---

## Explore Data Structure and Cleaning

The Oregon data contains 77069320 obsversations in total in year of 2013, which is a big data set. We want to figure out the structure for this data set and do some basic cleaning.

```{r, echo = F}
library(dplyr)
library(data.table)

# import data ------------------------------------------------------------------
read_path <- paste0("../data/gan_episodes_of_care.txt")

start_time <- Sys.time()
oregon_full <- fread(read_path, sep = "|", showProgress = T) # not change class
stop_time <-  Sys.time() - start_time 
# time it took
stop_time # 7.771713 mins

# Explore data structure
summary(as.factor(oregon_full$clmstatus))

# Clean data
oregon_full_df <- oregon_full %>%
  # filter non-missing value of gender, race 
  filter (gender!='U') %>%
  filter (is.na(race)==F)

# filter dx1, no ndc
oregon_full_dx1 <- oregon_full_df %>%
  filter (dx1!='*NULL*')

# filter ndc, no dx1
oregon_full_ndc <- oregon_full_df %>%
  filter (ndc!="")

fwrite(oregon_full_dx1, "../data/oregon_full_dx1.csv")
fwrite(oregon_full_ndc, "../data/oregon_full_ndc.csv")

write_path1 <- paste0('../data/','oregon_full_dx1.csv')
write_csv(oregon_full_dx1, write_path1)

write_path2 <- paste0('../data/','oregon_full_ndc.csv')
write_csv(oregon_full_ndc, write_path2)

```

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.