### 2017-02-16 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd
Compiling data exploring r markdown

---
title: "Oregon_Wildfire_reduced"
author: "Jingyang Liu"
date: "February 15, 2017"
output: html_document
---

# Explore Data Structure and Cleaning

The Oregon data contains 77069320 obsversations in total in year of 2013, which is a big data set. We want to figure out the structure for this data set and do some basic cleaning.

```{r}
library(dplyr)
library(data.table)
getwd()

# import data ------------------------------------------------------------------
# Read in first 2000000 lines for exploration.

read_path <- paste0("../data/gan_episodes_of_care.txt")
oregon_df <- fread(read_path, sep = "|", nrows = 200000, showProgress = T)
fwrite(oregon_df, "../data/oregon_epis_care_reduced.csv")

# I wrote a reduced dataframe of 2000,000 rows to get an idea of data structure

# import the reduced dataframe
start_time <- Sys.time()
oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",")
stop_time <-  Sys.time() - start_time

# time it took
stop_time # 1.968219 secs

```

## The overall summary  of data set
72 variables in total.
```{r overall summary}

# Explore data structure
summary(oregon_df)
```

## *personkey*, Unique person identifier
person's ID, maximum is 14381640
numeric; length: 8
```{r personkey}
summary(as.factor(oregon_df$personkey))
min(oregon_df$personkey)
max(oregon_df$personkey) # seems length is 8
```

## *year*, Calendar year in which claim was incurred
All is 2013
integer (numeric); length: 4
```{r year}
summary(as.factor(oregon_df$year)) # all 2013, 77069320
```

## *clmid*, Claim ID
numeric; length: 8
```{r claim id}
summary(as.factor(oregon_df$clmid))
```

## *line*, Claim line (Not Sure what this means)
numeric; range: 0-9999
```{r claim line}
unique(as.numeric(oregon_df$line)) # 1362 levels, 0 - 9999
```

## *clmtatus*, Claim status
D: 8676643 (11.26%), E: 3324496 (4.31%), P: 65068181 (84.43%)
character
```{r claim status}
summary(as.factor(oregon_df$clmstatus))
#       D        E        P
# 8676643  3324496 65068181

summary(as.factor(oregon_df$clmstatus))/77069320 # percentage
#          D          E          P
# 0.11258232 0.04313644 0.84428124
```

## *cob*, COB status (Coordination of benefits) 
blank: 23517978 (30.52%); N: 44438688 (57.66%); Y: 9112654 (11.82%)
character
``` {r cob}
summary(as.factor(oregon_df$cob)) 
#                 N        Y
# 23517978 44438688  9112654
summary(as.factor(oregon_df$cob))/77069320 # percentage
#                   N         Y
# 0.3051536 0.5766067 0.1182397
```

## *paytype*, Pay type
C (58.50%), D (31.77%), G (2.08%), P (3.53%), T (4.12%)
character
``` {r pay type}
summary(as.factor(oregon_df$paytype)) #C, D, G, P, T
summary(as.factor(oregon_df$paytype))/77069320 # percentage
#         C          D          G          P          T
# 0.58498296 0.31771349 0.02081743 0.03532560 0.04116052
```

## *prod*, Product code
CHP, etc. Insurance type/product code [web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=194604000)
character
```{r product code}
summary(as.factor(oregon_df$prod))
```

## *payer*, APAC payer 
All Payer All Claims [web](https://www.oregon.gov/oha/analytics/Pages/All-Payer-All-Claims.aspx)
character
```{r payer}
summary(as.factor(oregon_df$payer))
```

## *megflag*, Medical coverage flag
The medical coverage flag indicates whether this member is covered for medical expenses [web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=196401000)
0: (7.39%); 1: (91.57%); NA: (1.04%)
numeric; range: 0-1
```{r medical coverage flage}
summary(as.factor(oregon_df$medflag))
#       0        1     NA's
# 5697342 70573441   798537
summary(as.factor(oregon_df$medflag))/77069320 # percentage
#          0          1       NA's
# 0.07392490 0.91571381 0.01036128
```

## *rxflag*, Pharmacy coverage flag
This field indicates whether or not the member has prescription drug coverage. 
[web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=196401000)
0: (10.07%); 1: (88.89%); NA: (1.04%)
numeric; range: 0-1
```{r pharmacy converage flag}
summary(as.factor(oregon_df$rxflag))
#       0        1     NA's
# 7761348 68509435   798537
summary(as.factor(oregon_df$rxflag))/77069320 # percentage
#          0          1       NA's
# 0.10070607 0.88893265 0.01036128
```

## *ohvmhflag*, Oregon High value medical home flag
Required for participants in OHLC high value medical home initiative. 
[web](https://ushik.ahrq.gov/ViewItemDetails?&system=apcd&itemKey=194764000&enableAsynchronousLoading=true)
N: (69.15%); Y: (0.24%); NULL: (30.61%)
character
```{r Oregon High value medical home flag}
summary(as.factor(oregon_df$ohvmhflag))
#        N   *NULL*        Y
# 53291284 23589431   188605
summary(as.factor(oregon_df$ohvmhflag))/77069320 # percentage
#           N      *NULL*           Y
# 0.691472093 0.306080695 0.002447212
```

## *pebb*, PEBB flag (NOT Provided)
0: (95.95%); 1: (4.05%)
numeric; range: 0-1
```{r pebb flag}
summary(as.factor(oregon_df$pebb))
#        0        1
# 73949154  3120166
summary(as.factor(oregon_df$pebb))/77069320 # percentage
#          0          1
# 0.95951481 0.04048519
```

## *oebb*, OEBB flag (NOT Provided)
0: (96.39%); 1: (3.61%)
numeric; range: 0-1
```{r oebb flag}
summary(as.factor(oregon_full$oebb))
#        0        1
# 74285962  2783358
summary(as.factor(oregon_full$oebb))/77069320 # percentage
#          0          1
# 0.96388501 0.03611499
```

## *patid*, Encrypted patient ID
numeric
```{r Encrypted patient ID}
summary(as.factor(oregon_full$patid))
```

## *gender*, Gender
F: (60.56%); M: (39.38%); U (for unknown): (0.06%)
character
```{r gender}
summary(as.factor(oregon_full$gender))
#        F        M        U
# 46672307 30350771    46242
summary(as.factor(oregon_full$gender))/77069320 # percentage
#            F            M            U
# 0.6055886700 0.3938113247 0.0006000053
```

## *yob*, Birth year
old:1901 (116 years) young:2016 (1 year)
numeric
```{r birth year}
summary(as.factor(oregon_full$yob)) # old:1901 young:2016
```

## *race*, Race (Not sure the category means)
1:(31%); 2: (2%); 3: (0.8%); 4: (0.3%); 5: (0.1%); 6: (0.3%); 9: (65%); NA: (0.03%)
numeric; range: 1-9
```{r race}
summary(as.factor(oregon_full$race))
#        1        2        3        4        5        6        9     NA's
# 24132190  1611193   620854   203550    85465   248416 50141889    25763
summary(as.factor(oregon_full$race))/77069320 # percentage
#            1            2            3            4            5            6
# 0.3131231727 0.0209057638 0.0080557867 0.0026411288 0.0011089367 0.0032232800
#            9         NA's
# 0.6506076478 0.0003342835
```

## *ethn*, Ethnicity (Not sure the category means)
1:(0.76%); 2: (9.60%); 3: (89.61%); NA: (0.03%)
numeric; range: 1-3
```{r ethnicity}
summary(as.factor(oregon_full$ethn))
#        1        2        3     NA's
#  583948  7398053 69061589    25730
summary(as.factor(oregon_full$ethn))/77069320 # percentage
#            1            2            3         NA's
# 0.0075769191 0.0959921925 0.8960970332 0.0003338553
```

## *lang*, Primary spoken language 
Not sure the category means. This web [web](http://www.niso.org/topics/ccm/ccmstandards/) might help.
1:(47.70%); 2: (1.34%); 3: (0.001%); 4: (0.26%); 8: (0.65%); 9: (49.31%); NA: (0.74%)
numeric; range: 1-9
```{r language}
summary(as.factor(oregon_full$lang))
#        1        2        3        4        8        9     NA's
# 36761368  1031059      587   203152   501575 38002063   569516
round(summary(as.factor(oregon_full$lang))/77069320, 5)  # percentage
#       1       2       3       4       8       9    NA's
# 0.47699 0.01338 0.00001 0.00264 0.00651 0.49309 0.00739
```

## *MSA*, Metropolitan and Micropolitan Statistical Areas Main???
numeric
```{r MSA}
summary(as.factor(oregon_full$MSA))
```

## *STATE*, Standard two character abbreviation
OR: (99.33%); blank: (0.1%)
character
```{r State}
summary(as.factor(oregon_full$STATE)) # OR: 76555225; blank: 78646
76555225/77069320
```

## *ZIP*, Zip code
Freely available in the public domain. NA: (0.1%)
numeric
```{r ZIP code}
summary(as.factor(oregon_full$ZIP)) # NA: 78646
```

## *fromdate*, From date
YYYY-MM-DD
character
```{r from date}
summary(as.factor(oregon_full$fromdate))
unique(as.factor(oregon_full$fromdate))
```

## *todate*, To date
YYYY-MM-DD
character
```{r to date}
summary(as.factor(oregon_full$todate))
```

## *paydate*, Pay date
YYYY-MM-DD
character
```{r pay date}
summary(as.factor(oregon_full$paydate))
```

## *paid*, Total payment
NA: (0.36%)
numeric
```{r total payment}
summary(as.factor(oregon_full$paid)) # NA: 279811
```

## *tob*, Type of bill
blank: (52.61%); NULL: (25.12%)
character
```{r type of bill}
summary(as.factor(oregon_full$tob)) # blank: 40545960 NULL: 19358510 
```

## *pos*, Place of service code
Not sure. [web](https://www.cms.gov/Medicare/Coding/place-of-service-codes/Place_of_Service_Code_Set.html) migh help.
NULL: (9.22%)
character
```{r place of servis code}
summary(as.factor(oregon_full$pos)) # NULL: 7106719
```

## *dx1*, Principal diagnosis (For ICD9 code)
[web](https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html)
NULL: (31.81%)
character
```{r dx1}
summary(as.factor(oregon_full$dx1)) # NULL: 24513232
```

## *dx2*, Diagnosis 2
NULL: (59.31%)
character
``` {r dx2}
summary(as.factor(oregon_full$dx2)) # NULL: 45707452
```

## *dx3*, Diagnosis 3
NULL: (72.67%)
character
``` {r dx3}
summary(as.factor(oregon_full$dx3)) # NULL: 56005306
```

## *dx4*, Diagnosis 4
NULL: (81.46%)
character
``` {r dx4}
summary(as.factor(oregon_full$dx4)) # NULL: 62777450
```

## *dx5*, Diagnosis 5
NULL: (93.42%)
character
``` {r dx5}
summary(as.factor(oregon_full$dx5)) # NULL: 71995284
```

## *dx6*, Diagnosis 6
NULL: (95.23%)
character
``` {r dx6}
summary(as.factor(oregon_full$dx6)) # NULL: 73390409
```

## *dx7*, Diagnosis 7
NULL: (96.31%)
character
```{r dx7}
summary(as.factor(oregon_full$dx7)) # NULL: 74225796
```

## *dx8*, Diagnosis 8
NULL: (97.00%)
character
```{r dx8}
summary(as.factor(oregon_full$dx8)) # NULL: 74759988
```

## *dx9*, Diagnosis 9
NULL: (98.55%)
character
```{r dx9}
summary(as.factor(oregon_full$dx9)) # NULL: 75949570
```

## *dx10*, Diagnosis 10
NULL: (98.81%)
character
```{r dx10}
summary(as.factor(oregon_full$dx10)) # NULL: 76151124
```

## *dx11*, Diagnosis 11
blank: (98.35%)
character
``` {r dx11}
summary(as.factor(oregon_full$dx11)) # blank: 75801079
```

## *dx12*, Diagnosis 12
blank: (98.61%)
character
```{r dx12}
summary(as.factor(oregon_full$dx12)) # blank: 76000733
```

## *dx13*, Diagnosis 13
blank: (98.86%)
character
```{r dx13}
summary(as.factor(oregon_full$dx13)) # blank: 76193335
```

## *poa1*, POA code 1
N: (0.07%); Y: (5.94%); blank: (93.99%)
character
```{r poa1} 
summary(as.factor(oregon_full$poa1))
#                 N        Y
# 72434455    58842  4576023
summary(as.factor(oregon_full$poa1))/77069320 # percentage
#                         N            Y
# 0.9398610887 0.0007634945 0.0593754168
```

## *poa2*, POA code 2
N: (0.34%); Y: (3.15%); blank: (96.51%)
character
```{r poa2}
summary(as.factor(oregon_full$poa2))
#                 N        Y
# 74380115   261029  2428176
summary(as.factor(oregon_full$poa2))/77069320 # percentage
#                       N           Y
# 0.965106673 0.003386938 0.031506389
```

## *poa3*, POA code 3
N: (0.34%); Y: (2.75%); blank: (96.91%)
character
```{r poa3}
summary(as.factor(oregon_full$poa3))
#                 N        Y
# 74688528   264287  2116505
summary(as.factor(oregon_full$poa3))/77069320 # percentage
#                       N           Y
# 0.969108434 0.003429212 0.027462355
```

## *px1*, Principal inpt procedure
blank: (52.61%); NULL: (45.38%)
character
```{r px1}
summary(as.factor(oregon_full$px1)) # blank: 40545960; NULL: 34975994
```

## *px2*, Procedure 2
blank: (52.61%); NULL: (46.06%)
character
```{r px2}
summary(as.factor(oregon_full$px2)) # blank: 40545960 NULL: 35502262
```

## *px3*, Procedure 3
blank: (52.61%); NULL: (45.38%)
character
```{r px3}
summary(as.factor(oregon_full$px3)) # blank: 40545960; NULL: 34975994
```

## *proccode*, CPT or HCPCS procedure code
Current Procedural Terminology (CPT) Healthcare Common Procedure Coding System (HCPCS) 
[web1](https://www.cms.gov/Medicare/Coding/MedHCPCSGenInfo/index.html?redirect=/medhcpcsgenin)
[web2](https://www.cms.gov/Medicare/Coding/HCPCSReleaseCodeSets/index.html?redirect=/hcpcsrelea)
NULL: (34.88%)
character
```{r procedure ccode}
summary(as.factor(oregon_full$proccode)) # NULL: 26880271
```

```{r}
summary(as.factor(oregon_full$mod1))
```

```{r}
summary(as.factor(oregon_full$mod2))
```

```{r}
summary(as.factor(oregon_full$megcode))
```

```{r}
summary(as.factor(oregon_full$megdesc))
```

```{r}
summary(as.factor(oregon_full$megbodysys))
```

```{r}
summary(as.factor(oregon_full$megstage))
```

```{r}
summary(as.factor(oregon_full$megtype))
#                   9 NOT-CATEGORIZED             Acute           Chronic
#              1751           6979372          34062467          31197648
#         Well Care
#           4828082
```

```{r}
summary(as.factor(oregon_full$megcomplete))
#        0        1     NA's
# 10090656 66976913     1751
```

```{r}
summary(as.factor(oregon_full$megnum))
```

```{r}
summary(as.factor(oregon_full$megdays))
```

```{r}
summary(as.factor(oregon_full$megprorate))
```

```{r}
summary(as.factor(oregon_full$megoutlier))
#        0        1
# 54675387 22393933
```

```{r}
summary(as.factor(oregon_full$meglow))
#    FALSE     TRUE
# 75145616  1923704
```

```{r}
summary(as.factor(oregon_full$meghigh))
#    FALSE     TRUE
# 56599091 20470229
```

``` {r}
# ndc
summary(as.numeric(oregon_full$ndc)) # blank: 53599072
```

```{r}
summary(as.factor(oregon_full$rxclass))
```

```{r}
summary(as.factor(as.character(oregon_full$qtydisp)))
```

```{r}
summary(as.factor(oregon_full$rxdays))
```

```{r}
summary(as.factor(oregon_full$daw))
```

```{r}
summary(as.factor(oregon_full$ptstatus))
```

```{r}
summary(as.factor(oregon_full$los))
```

```{r}
summary(as.factor(oregon_full$msdrg))
```

```{r}
summary(as.factor(oregon_full$icd_ver))
#    FALSE     NA's
# 52556088 24513232
```

## Clean data

```{r }
# Clean data, 76986701 left (maybe later)
#oregon_full_df <- oregon_full %>%
# filter non-missing value of year of birth, gender, race, STATE
#  filter (is.na(yob)==F) %>%
#  filter (gender!='U') %>%
#  filter (is.na(race)==F) %>%
#  filter (STATE!='')

# Explore no dx1 and no ndc, see if the dx2 or other reasons, 1042984 obs
# oregon_full_miss <- oregon_full %>%
#  filter (dx1=='*NULL*') %>%
#  filter (ndc=="")

# fwrite(oregon_full_miss, "../data/oregon_full_miss.csv")

# read_path1 <- paste0("./data/oregon_full_miss.csv")
# oregon_miss <- fread(read_path1, sep = ",")

# oregon_miss_df <- oregon_miss %>%
#   filter (dx2!='*NULL*') # 981 obs

# filter dx1, no ndc, 52490460
# oregon_full_dx1 <- oregon_full_df %>%
#   filter (dx1!='*NULL*')

# filter ndc, no dx1, 23453379
# oregon_full_ndc <- oregon_full_df %>%
#  filter (ndc!="")

# fwrite(oregon_full_dx1, "../data/oregon_full_dx1.csv")
# fwrite(oregon_full_ndc, "../data/oregon_full_ndc.csv")

```