### 2017-03-24 r_scripts/parallel_presentation.Rmd
Update parallel_presentation.Rmd

---
title: "Parallel for Oregon Zip and WRF_Grid Intersection Proportion"
author: "Jingyang Liu"
date: "March 23, 2017"
output: slidy_presentation
---

```{r library setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rgdal) # package for shape files
library(sp)
library(rgeos) # rgeos package contains the intersect and area commands I need
library(dplyr)
library(data.table)
library(readxl)
# parallel computing libraries
# library(foreach) 
library(doParallel)
library(parallel)
library(readr)
```

```{r shapefile prepare, include=FALSE}
## Imoport WRF Grid file of Oregon
setwd("C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire")
grid_dir <- paste0('./shapefile/oregon_grid.shp')
smoke_grid <- readOGR(dsn = grid_dir, layer = 'oregon_grid') # wrf_grid

## Import U.S. shapefile
shp_dir <- paste0('./shapefile/tl_2013_us_zcta510/tl_2013_us_zcta510.shp')
us_zip_2013 <- readOGR(dsn = shp_dir, layer = 'tl_2013_us_zcta510')

## Filter the U.S. shapefile only to Oregon State Zip Code
read_path2 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire/', 
                     'oregon_zipcode.csv')
or_zip <- read_csv(read_path2)
names(or_zip) <- c('zip','type','city','county','area')
or_zip_map <- us_zip_2013[us_zip_2013$ZCTA5CE10 %in% or_zip$zip,]
or_zip <- as.character(sort(or_zip_map@data$ZCTA5CE10))

nad83 <- '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
proj4string(smoke_grid) <- nad83
```

## Shapefile preparation: 2013 Oregon Zip code Shapefile and WRF Grid
The shapefile of Oregon can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
The whole U.S shapefile can be found from United States Census website:
<br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas
<br/>The Oregon zip code can also be found from: 
<br/>https://www.unitedstateszipcodes.org/or/

The plot below shows a Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.
```{r overlay}
plot(or_zip_map)
plot(smoke_grid)
plot(or_zip_map, add=T)
```

The Oregon shapefile contains some lack or extra, which may because the zip code changes (but U.S. shapefile is for 2013).

## Steps
Choose zip code = 94705. WRF Grid = 261 and 262.
```{r test}
test_zip <- c(97405)
test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]

plot(test_zip_map)
plot(smoke_grid, add = T)
invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
               labels=as.character(smoke_grid$WRFGRID_ID)))

```

```{r}
test_grid <- over(smoke_grid, test_zip_map)
summary(test_grid) # 14 grids over the zip code
test_grid2 <- test_grid %>% filter(!is.na(ZCTA5CE10))
test_grid2

# test zip map over zip grid
plot(smoke_grid)
plot(test_zip_map, add = T)
zip_over_grid <- over(test_zip_map, smoke_grid)
summary(zip_over_grid) # this way retains the values

shape_zip <- SpatialPolygons(test_zip_map@polygons)
shape_grid <- SpatialPolygons(smoke_grid@polygons)
plot(shape_zip)
plot(shape_grid, add = T)
# try and plot the values for this zipcode for 9/21
invisible(text(getSpPPolygonsLabptSlots(smoke_grid), 
               labels=as.character(smoke_grid$WRFGRID_ID)))

# output the WRF Grid 261 and 262 for test
wrf_grid_261 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 261, ]
plot(wrf_grid_261)

wrf_grid_262 <- smoke_grid[smoke_grid@data$WRFGRID_ID == 262, ]
plot(wrf_grid_262)

# calculate the area of grid 297
gArea(SpatialPolygons(wrf_grid_261@polygons))
gArea(SpatialPolygons(wrf_grid_262@polygons))

# Subset to one zip code and two different WRF grids ---------------------------
# plot zip and 2 grids
plot(test_zip_map)
#invisible(text(getSpPPolygonsLabptSlots(test_zip_map), 
#               labels=as.character(test_zip_map$ZCTA5CE10)))
plot(wrf_grid_261, add = T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
               labels=as.character(wrf_grid_261$WRFGRID_ID)))
plot(wrf_grid_262, add=T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_262), 
               labels=as.character(wrf_grid_262$WRFGRID_ID)))

# Start with the intersection with wrf grid 297 and zipcode
# first I need to convert the spatial polygon to just polygon
poly_261 <- SpatialPolygons(wrf_grid_261@polygons)
shape_zip <- SpatialPolygons(test_zip_map@polygons)

zip_261_int <- gIntersection(poly_261, shape_zip)

plot(zip_261_int)
prop_261_int <- gArea(zip_261_int)/gArea(poly_261)
prop_261_int # 41.8% of grid is covered by zip

# now what about grid 296
poly_262 <- SpatialPolygons(wrf_grid_262@polygons)
shape_zip <- SpatialPolygons(test_zip_map@polygons)

zip_262_int <- gIntersection(poly_262, shape_zip)

plot(zip_262_int)
prop_int_262 <- gArea(zip_296_int)/gArea(poly_262)
prop_int_262 # 6.0% of grid is covered by zip

```

## Why using for loop?

<ul class="incremental"> 
  <li>One choice: <br/>  simple calculation; copy and paste</li> 
  <li>Another choice: <br/>  for loop</li> 
</ul>

## Apply Family

- The apply() Function
- The lapply() Function
- The sapply() Function
- The tapply() Function
- The mapply() Function

## Parallel on server

```{r cars, echo = TRUE}
summary(cars)
```