### 2017-04-17 r_scripts/analysis/or_result_zip_smoke.Rmd
Create or_result_zip_smoke.Rmd

---
title: "Oregon: Zip-Level Wildfire Smoke and Health Outcomes (open in wildfire, not script)"
author: "Jingyang Liu"
date: "April 7, 2017"
output: html_document
---

## Overview

This document contains results for the association between wildfire smoke PM~2.5~ and cardiovascular (CVD) and respiratory health outcomes. This particular markdown document focuses on the zip code-level analysis. Herein, I compare different PM~2.5~ estimation methods described in further detail below.

```{r include=FALSE}
library(tidyverse)
library(data.table)
library(survival)
library(htmlTable)
library(readr)
```

```{r data import, include = F, eval = FALSE}

### Limit to only ER visits
## pos = 23
var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')

N <- NA
n <- 1

for (i in var_list){
  read_path <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
  or_disease <- read_csv(read_path)

  or_disease <- or_disease %>% 
    filter(pos=="23") %>%
    # filter the unknown gender
    filter(gender!= "U")

  # Create a permanent case-cross over dataset
  file_name <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")

  # write permanent dataset
  write_csv(or_disease, paste0(file_name))
  N[n] = length(unique(or_disease$personkey))
  n = n+1
}

num_person <- data.frame(rbind(var_list, N))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
## Analysis
path <- paste0("../data_new/update/ER")

resp_casecross <- read_csv(paste(path, 
                                 "or_respiratory_may_to_sep_time_strat_casecross_er.csv", 
                                 sep = "/"))

# asthma
asthma_casecross <- read_csv(paste(path, 
                                   "or_asthma_may_to_sep_time_strat_casecross_er.csv", 
                                   sep = "/"))

# copd 
copd_casecross <- read_csv(paste(path, 
                                 "or_copd_may_to_sep_time_strat_casecross_er.csv", 
                                 sep = "/"))

# pneum or bronchitis
pneum_casecross <- read_csv(paste(path, 
                                  "or_pneumonia_may_to_sep_time_strat_casecross_er.csv",
                                  sep="/"))
 # acute bronchitis
acute_bronch_casecross <- read_csv(paste(path, 
                                         "or_acute_bronch_may_to_sep_time_strat_casecross_er.csv",
                                         sep = "/"))
# cvd
cvd_casecross <- read_csv(paste(path, 
                                "or_cvd_may_to_sep_time_strat_casecross_er.csv",
                                sep="/"))
# arrhythmia
arrhythmia_casecross <- read_csv(paste(path, 
                                       "or_arrhythmia_may_to_sep_time_strat_casecross_er.csv",
                                       sep="/"))
# cerebral vascular
cereb_vas_casecross <- read_csv(paste(path, 
                                      "or_cerbrovas_dis_may_to_sep_time_strat_casecross_er.csv",
                                      sep="/"))
# heart failure
hf_casecross <- read_csv(paste(path, 
                               "or_heart_failure_may_to_sep_time_strat_casecross_er.csv", 
                               sep="/"))
# ischemic heart disease
ihd_casecross <- read_csv(paste(path, 
                                "or_isch_heart_dis_may_to_sep_time_strat_casecross_er.csv",
                                sep="/"))
# myo infarc
mi_casecross <- read_csv(paste(path, 
                               "or_myocardial_infarc_may_to_sep_time_strat_casecross_er.csv", sep="/"))

# broken arm
broken_arm_casecross <- read_csv(paste(path, 
                                       "or_broken_arm_may_to_sep_time_strat_casecross_er.csv",
                                       sep="/"))

# read zip pm for counts by zip
smoke_path <- paste0("../data/Oregon/")
zip_pm <- read_csv(paste0(smoke_path, "zip_pm_to_merge_with_chars.csv"))
```

## Methods Description

In these comparisons, we examine various methods of smoke/PM~2.5~ estimations and associations with health outcomes using a time-stratified case-crossover study design. Health outcomes for a patient with a primary diagnosis of cardiopulmonary health outcomes and their date of admission  (index time) were identified. We then created counterfactual observations for each patient for the same day of the week for the entire wildfire season (MAy 1st to Sep 30th, 2013). We further limited our analyses to claims that were coded as emergency or urgent visits to eliminate bias from patients going in for elective/planned procedures. However, we found the filtered population is much larger than the population we chose in Washington, so we choose only the emegency room (hospital) visits. For patients who have a index time closer to May 1, 2013 or closer to September 30, 2013, their referent observations before or after these dates will be excluded as I will not be able to assign estimates of PM~2.5~ to these referent observations. However, this is not a big issue for a time-stratified design as other referent values throughout the time period average out.

The **health outcomes** of interest are all respiratory, asthma, COPD, pneumonia, acute bronchitis, cardiovascular disease, heart failure, ischemic heart disease, and myocardial infarction. I also included broken arm, which I hypothesize to not be associated with wildfire smoke exposure and use as a check.

As for **exposure methods**, there are three main estimation methods for PM~2.5~: WRF-Chem Smoke (which substracts the WRF-Chem no fire emission from WRF-chem). For the WRF-Chem variable with the 'smoke' designator, this is WRF-Chem - WRF-Chem no fire. For Geo-Weighted Regression, and Kriging with 'smk' designator, I subtracted off the 'Background' estimates of smoke, which I believe are the monthly averages of PM~2.5~ for a given grid. However, the 'Background' is the mean of total period from May 1st to Sep 30th 2013 so I have to assign the mean for every day.

The **analytic method** is the conditional logistic regression model using the *survival* package in R. Each conditional logistic regression model accounts for the subject, and adjusts for temperature from the WRF-Chem model. The conditional logistic regression model \beta represents the change in risk of an event associated with a short-term unit increase in exposure, and can be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times in the referent window. See Janes et al. 2005, **Statistics in Medicine**.

I've run a couple different scenarios for selecting referent periods, each of which has some paper that may justify the benefit of the design, relative to other methods.

## Population-Weighted PM~2.5~ Assigned by Zip Code

### Descriptives

Note for following tables: I wrote out 'Greater than rather' than using  '>' as it was increasing the following font and looked weird.

```{r descriptive table, echo = F, warning = F, results='asis'}

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
  pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
  cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
  broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 'Kriging Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# set list and do cross tabs to find values
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  i = 2
# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- df_to_loop %>%
    # only look at outcomes
    filter(outcome == 1) %>%
    # add another row that makes sure there is a person <15 in the dataframe
    # tricking xtabs to produce a 0 cell for the outcome for age <15
    add_row(outcome = 0, age_ind = 0)

  # cross tabs
  outcome_n <- xtabs(~ outcome, df_analysis)
  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
  # empty matrix
  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)

  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
                                 "age_65", "female", "male")

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name
  # fill n
  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
  # age <15
  point_estimates[, 3] <- cross_tab_age[2, 1]
  # age 15 to 65
  point_estimates[, 4] <- cross_tab_age[2, 2]
  # age >65
  point_estimates[, 5] <- cross_tab_age[2, 3]
  # male == 0
  point_estimates[, 7] <- cross_tab_sex[1, 1]
  # female == 1
  point_estimates[, 6] <- cross_tab_sex[1, 2]

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist) %>% 
  # find proportions/percents for each strata in a row
  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
         age_65_pr, female, female_pr, male, male_pr)

#str(combined_point_est_df)

tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
           caption = "Number of cases for each outcome observed from July 1st to October 31st, 2012",
           # column headers
           header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
                      "(%)", "Female", "(%)", "Male", "(%)"),
           # column spanner
           cgroup = c("","Age Category", "Sex"), 
           n.cgroup = c(2, 6, 4),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llccccc" # column alignment,
            ) # end table

print(tab)

```

### Time-Stratified Case-Crossover (referents same day of week within fire season July - October)

For these analyses and for the final results in the paper, I used referents within the entire wildfire season (July to October). This time-stratified referent strategy allows even more referent observations to average to 'smooth' out some of those extreme PM~2.5~ concentrations in some referent periods. For my sensitivity analysis of varying referent periods, please see other scripts.

Adjusting for CDC Met Variables: Rish has provided zip-level meteorologic data on wind speed, precipitation, relative humidity, and temperature. Note a slight difference that will likely need to be addressed in discussion is that these measures are weighted based on census tract, not on the WRF grid like our measures of smoke.

```{r echo = F, warning = F, results='asis'}

df_list <- list(resp_casecross, asthma_casecross, copd_casecross, 
                pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  which(colnames(df_to_loop)=="background_pm_zip") # code to find column numbers, 91
  pm_estimates_df <- df_to_loop[, c(88, 93, 92, 94)]/10  # create 10 unit increases

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_zip)) %>%

    # join in cdc met data
    # full_join(daily_zip_met, by = c("ZIPCODE", "date")) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # empty df for table
  table_df <- data.frame()

  # empty matrix
  point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name

  # second loop to run a model for each pm estimation method
  for(j in 39:41){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # some model/DAG checks of data rish provided

    #glimpse(df_analysis)

    # adjusting for admit month is not possible as it's the same var for the subject.
    # adjusting for month smoke is a collider and should not be adjusted for.
    #summary(clogit(outcome ~ geo_smk_pm_zip + daily_meanRH + daily_meanT + daily_meanWS + daily_meanPrcp +
    #                as.factor(month_smk) + strata(PATIENTID), df_analysis))
    #outocme association: daily_meanRH,

    # check for confoudning following DAG assumptions
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    #summary(clogit(outcome ~ wrf_temp_zip + strata(PATIENTID), df_analysis))
    # smoke predictor association: daily_meanRH, 
    #summary(lmer(geo_smk_pm_zip ~ daily_meanWS + (1 | PATIENTID), df_analysis ))
    #summary(lmer(wrf_temp_zip ~ daily_meanT + (1 | PATIENTID), df_analysis ))
    #cor(df_analysis$wrf_temp_county, df_analysis$daily_meanT)

    # populate matrix
    row_n <- j-38

    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  }

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2, 3:7)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 and Health Outcomes",
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "Obs.", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for CDC temperature, relatively humidity, wind speed, and precipitation; accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
combined_point_est_df$outcome <- factor(combined_point_est_df$outcome, 
                                        levels = unique(combined_point_est_df$outcome))

combined_point_est_df$pm_method <- factor(combined_point_est_df$pm_method, 
                                          levels = unique(combined_point_est_df$pm_method))

combined_point_est_df$n <- as.numeric(combined_point_est_df$n)
combined_point_est_df$n_events <- as.numeric(combined_point_est_df$n_events)
combined_point_est_df$odds_ratio <- as.numeric(combined_point_est_df$odds_ratio)
combined_point_est_df$lower95 <- as.numeric(combined_point_est_df$lower95)
combined_point_est_df$upper95 <- as.numeric(combined_point_est_df$upper95)
combined_point_est_df$se <- as.numeric(combined_point_est_df$se)
combined_point_est_df$p_val <- as.numeric(combined_point_est_df$p_val)

# subset out global method
combined_point_est_df <- combined_point_est_df

## ggplot
print_plot <- ggplot(combined_point_est_df, 
                     aes(x = pm_method, y = odds_ratio, colour = pm_method)) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  geom_point() + #geom_text(vjust = 0, nudge_x = 0.3) +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome, nrow = 3) +
  geom_hline(yintercept = 1, linetype=2) +
  #    ggtitle('Association Between PM2.5 \n from Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  #xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        #strip.text = element_text(size = 8),
        # axis element
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("./fig_only_er.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")
```

#### Result Summary: Time-Stratified Case-Crossover Design (same day of week within the season)

A general interpretation is as follows:
A 10 $\mu$g/m^3^ increase in PM~2.5~ attributed to wildfire smoke was associated with a X% increase in the risk for a health event. Where X% is the odds ratio - 1 (e.g. OR of 1.20 - 1 = 20% increase). 
Note that the odds ratio approximates the relative risk in this case due to the rare outcome assumption in that the calcuation for an odds ratio is similar to risk ratio calcuation.

We see that all exposure methods are significantly associated with an increase in all respiratory outcomes and asthma. The WRF-Chem smoke method is associated with cerebrovascular disease as well. As for the other models, Kriging, Global Regression, and Geo-Weighted Regression, those are associated with COPD. I believe this is the most justifiable design approach that would average any extreme values of PM~2.5~ across the wildfire season for the vector of referent values for a particular subject.