### 2017-05-10 r_scripts/analysis/county_timeseries.Rmd
Create county_timeseries.Rmd

---
title: "Oregon 2013 Time-Series Analysis"
author: "Jingyang Liu"
date: "May 10, 2017"
output: html_document
---

```{r }
library(tidyverse)
library(readr)
library(lme4)
library(broom)
library(htmlTable)

# load time series dataframe
or_ts_df <- read_csv("../data_new/county_data/or_county_timeseries.csv")

# analysis dataframe
or_may_sep_df <- or_ts_df %>% 
  # create day of the week variable
  mutate(day = as.factor(weekdays(date))) %>% 
  # set missing outcome values to 0 since missing indicates no ER or urgent care 
  # visits on that date (which is reasonable in sparsely populated counties)
  mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:broken_arm_n) %>% 
  # create a binary smoke indicator variable based smk variables >5, >10, and >15 units 
  mutate(wrf_smk5 = ifelse(wrf_smk_pm >= 5, 1, 0), 
         wrf_smk10 = ifelse(wrf_smk_pm >= 10, 1, 0),
         wrf_smk15 = ifelse(wrf_smk_pm >= 15, 1, 0), 
         geo_smk5 = ifelse(geo_smk_pm >= 5, 1, 0),
         geo_smk10 = ifelse(geo_smk_pm >= 10, 1, 0),
         geo_smk15 = ifelse(geo_smk_pm >= 15, 1, 0),
         krig_smk5 = ifelse(krig_smk_pm >= 5, 1, 0),
         krig_smk10 = ifelse(krig_smk_pm >= 10, 1, 0),
         krig_smk15 = ifelse(krig_smk_pm >= 15, 1, 0),
         season = ifelse(date >= "2013-05-01" & date <= "2013-06-21", "spring", 
                  ifelse(date >= "2013-06-22" & date <= "2013-09-22", "summer",
                  ifelse(date >= "2013-09-23" & date <= "2013-12-21", "fall",
                                "other")))) %>% 
  # rescale pm variables to 10 units
  mutate_each(funs(unit10 = ./10), wrf_pm:krig_smk_pm)
  # select(1:2, 31:33, 36, 34:35, 38:44, 28:29, 30, 55, 58, 
  #        65, 64, 63) # subset and order variables used in analysis

# outcomes to loop through
outcome_list <- c('All ER', 'All Respiratory', 'Asthma', 'Pneumonia', 
                  'Acute Bronchitis','COPD', 'Cardiovascular Disease',
                  'Ischemic Heart Disease', 'Arrhythmia', 'Heart Failure', 
                  'Cerebrovascular Disease', 'Myocardial Infarction',
                  'Broken Arm')

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('Geo-Weighted Smoke', 'Kriging Smoke', 'WRF-Chem Smoke')

# create an empty list to row bind dataframes together
datalist <- list()

#glimpse(wash_jul_oct_df)
# finding columns to retain
#which(colnames(wash_jul_oct_df)=='geo_smk_pm_unit10')
#summary(wash_jul_oct_df)
# loop through outcomes
for(i in 6:18){
  i = 8
  # empty matrix for table
  point_estimates <- matrix(nrow = 3, ncol = 5, byrow = T)

  colnames(point_estimates) <- c('outcome', 'pm_method', 'risk_ratio', 
                                 'lower95', 'upper95')

  # variable to model 
  outcome_name <- outcome_list[i-5]
  # fill outcome name
  point_estimates[, 1] <- outcome_name

  # loop through pm estimation methods
  for(j in 25:27){
    j = 25
    var_name <- method_list[j-22]

    # random effects model where county is treated as a random intercept 
    mixed_mod_adj <- glmer(or_may_sep_df[[i]] ~ or_may_sep_df[[j]] + day + season +
                             wrf_temp + (1|county), or_may_sep_df, family = "poisson")

    model_df <-tidy(mixed_mod_adj)

    # populate matrix
    row_n <- j-22
    # method
    point_estimates[row_n, 2] <- var_name
    # rate ratio
    point_estimates[row_n, 3] <- round(exp(model_df[2,2]), 3)
    # 95% lower bound
    point_estimates[row_n, 4] <- round(exp((model_df[2,2]) -
                                             1.96*(model_df[2,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 5] <- round(exp((model_df[2,2]) +
                                             1.96*(model_df[2,3])), 3)

    # save point estimates as a dataframe
    point_est_df <- as_data_frame(point_estimates)

  } # end methods loop

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df

} # end outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist)

# subset columns I want to put in to the table
table_df <- combined_point_est_df %>% select(2:5)

tab <- htmlTable(txtRound(table_df, digits = 3, 1:3), 
                 caption = paste0("County-Level Time-Series: Association between",
                                  " a 10 ug/m^3 in PM2.5 and Health Outcomes"),
                 # row group by outcome
                 rgroup = outcome_list,
                 n.rgroup = c(rep(3, 12)), # 4 rows for each method for each outcome
                 # column headers
                 header = c("Method", "RR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(2, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc", # column alignment,
                 tfoot="&dagger; Adjusted for temperature, day of the week, and season."
) # end table

print(tab)