### 2017-11-03 r_scripts/analysis/oregon_medication_ts.Rmd

---
title: "Oregon 2013 Medication for Inhalers Time Series and Distributed Lag"
author: "Jingyang Liu"
date: "May 17, 2017"
output: html_document
---

```{r library, include=FALSE, echo=FALSE}
library(tidyverse)
# library(readxl) 
library(survival)
library(htmlTable)
library(knitr)
library(broom)
library(lme4)

# two-stage distributed lag model libraries
# library(dlnm)
# library(mvmeta)
# library(splines)

```

### Data Import

```{r import strat casecrossover data, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
asthma_ndc <- read_csv(read_path6)
# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

```

## County-level time series

### County Time series plots

```{r prepare pop and pm, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Join data with names of Oregon counties 
oregon_fips <- read_csv('../instructions/oregon_FIPS.csv')
# remove the "County" character
# factor(oregon_fips$county)
oregon_fips$county <- gsub(" County", "", as.character(factor(oregon_fips$county)))

oregon_fips <- oregon_fips %>%
  mutate(st_county_fips = with(oregon_fips, paste0(st_code, fips)))

### Join county with zip
read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
or_zip_county <- read_csv(read_path)

or_zip_county <- or_zip_county %>%
  select(zip, county_name) %>%
  rename(county = county_name) %>%
  rename(ZIP = zip) %>%
  full_join(oregon_fips, by = "county") %>%
  select(ZIP, county, state, st_county_fips) %>%
  rename(fips = st_county_fips)

or_zip_county$county[which(or_zip_county$county=="Hood.River")] <- "Hood River"

read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
or_pop <- read_csv(read_path9)

or_pop <- or_pop %>%
  rename(county = GEO_display_label, pop = respop72013) %>%
  select(county, pop) 
or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))

```

### Prepare

1. Make 10 units geo smoke
2. Make county population (to calculate rate)
3. Make weekend/weekday variables

```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
read_path7 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
or_ndc_county_ts <- read_csv(read_path7)

or_ndc_county_ts <- or_ndc_county_ts %>%
  left_join(or_pop, by = "county") %>%
  mutate(geo_smk_pm10 = 10*geo_smk_pm) %>%
  mutate(rate_per_100k = (n_obs/pop)*100000) %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA)))

or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-05-27")] <- "weekend"
or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-07-04")] <- "weekend"
or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-09-02")] <- "weekend"

plot_ts <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
           geom_jitter()

print(plot_ts)

# save figure
ggsave("../plot_new/ndc_rate_ts.pdf", plot = plot_ts, 
       width = 12, height = 8, units = "in")

plot_county <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
               geom_jitter() + facet_wrap(~county)

plot_county_line <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
                    facet_wrap(~county) + geom_line()

print(plot_county)
print(plot_county_line)
# save figure
ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
       width = 12, height = 8, units = "in")
ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
       width = 12, height = 8, units = "in")

### check if the time series has the association with get smoke
a <- glm(rate_per_100k ~ geo_smk_pm10, data = or_ndc_county_ts, family = "poisson")
summary(a)

plot_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
  geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
  # scale_colour_manual(values = c("red", "blue")) +
  # truncated y axis
  #scale_y_continuous(limits = c(0, 1.5)) +
  # facet_wrap(~county, scale  = "free") +
  ylab("Log Outcome Rate per 100,000 Persons") +
  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
  theme_bw()

plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
  geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
  # scale_colour_manual(values = c("red", "blue")) +
  # truncated y axis
  #scale_y_continuous(limits = c(0, 1.5)) +
  facet_wrap(~county, scale  = "free") +
  ylab("Log Outcome Rate per 100,000 Persons") +
  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
  theme_bw()

plot(plot_health_smk)
plot(plot_county_health_smk)

ggsave("../plot_new/ndc_rate_health_smk.pdf", plot = plot_health_smk, 
       width = 12, height = 8, units = "in")

ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
       width = 12, height = 8, units = "in")

```

### Standard Regression Model

Same-day estimates of association between smoke via GWR and outcomes. County treated as a fixed effect covariate.

1. Quasipoisson

```{r regression model q, message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="quasipoisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_df[1,2:4] <- rr

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

rr_df

plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_rr.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

```

2. Poisson

```{r regression model,  message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_df[1,2:4] <- rr

rr_df

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_rr_poi.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

```

### Mixed models

1. Poisson

```{r mixed model,  message=FALSE, echo=FALSE}
# create empty dataframe
rr_mix_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_mix_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_mix_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glmer(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp + (1|county) + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_mix_df[1,2:4] <- rr

rr_mix_df

plot_mix_rr <- ggplot(rr_mix_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_mix_rr_poi.pdf", plot = plot_mix_rr, 
       width = 12, height = 8, units = "in")

```

```{r, eval = FALSE, include=FALSE}
asthma_ndc_df <- asthma_ndc_df %>%
  rename(ZIP = ZIPCODE) %>%
  left_join(or_zip_county, by = "ZIP") %>% # NA: 7005
  left_join(or_pop, by = "county")

```

```{r, eval=FALSE, include=FALSE}

# Import population weighted county PM2.5
pm_path <- paste0("../data_new/county_data/or_county_pop_wt_pm.csv")
county_pm <- read_csv(pm_path)

# create a working dataset 'or_2013'
or_ndc_county_time_series <- or_ndc_asthma %>% 
  # join in the washington county name based on CHARS county identifier
  left_join(or_zip_county, by = "ZIP") %>% 
  # rename 'admit_date_impute' to 'date'
  rename(date = dates) %>%
  # rename(date = admit_date_impute) %>% 
  # group by county of residence and date
  group_by(county, date) %>% 
  # sum up each primary diagnosis for each outcome for each day for each county
  summarise(n_obs = n()) %>% 
  # join daily county estimates of pm2.5
  left_join(county_pm, by = c('date', 'county'))

# which(is.na(or_ndc_county_time_series$county))
# or_ndc_county_time_series[113,]
# which(or_zip_county$ZIP=="97078")

# summary(or_ndc_asthma)
# summary(or_ndc_county_time_series)
# summary(as.factor(or_ndc_county_time_series$county)) # 125 is missing out of 3592 (3.16%)

# remove NA values
# which(is.na(or_ndc_county_time_series$county)) # 3468-3592
or_ndc_county_time_series <- or_ndc_county_time_series[-c(3468:3592),]

# write permanent dataframe

write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
write_csv(or_ndc_county_time_series, write_path5)

## Check for missing values in sparsely populated counties. set to 0 (But no)
# check_missing <- filter(or_ndc_county_time_series, is.na(n_obs))
# mutate each to fill in missing values (Still no)
# check2 <- check_missing %>% mutate_each(funs(new = ifelse(is.na(.), 0, .)), 
#                                         n_obs:wrf_temp)

```

### Time series of weekday and weekend in larger counties.
We check the time series for Multnomah, and it indeed shows more visits of weekdays than that in weekends (except some holidays such like Labor Day, Independent Day, etc.) We may choose the new covariates for weekday and weekend.

```{r echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
or_ndc_county_time_series_week <- or_ndc_county_time_series%>%
  mutate(day_admit = as.factor(weekdays(date)),
         month_admit = month(date)) %>%
  mutate(outcome = 1)

# Also check for Multnomah
multnomah_df <- or_ndc_county_time_series_week %>%
  filter(county == "Multnomah")

plot_mult_week <- ggplot(multnomah_df, aes(y =n_obs, x = date)) +
           geom_jitter() + facet_wrap(~day_admit)

print(plot_mult_week)

ggsave("../plot_new/multnomah_ts_week.pdf", plot = plot_mult_week, 
       width = 12, height = 8, units = "in")
```

### Time series of smoke exposure
The plot shows the times series of smoke PM 2.5 exposure. It seems that at the beginning of August or at the end of July the smoke has a peak. Also during the August the smoke is very high. The green one shows the WRF-Chem method; The bule one shows the Krigging method; And the red one, which we focus on, shows the GWR method.

```{r echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}

plot_smoke_ts_wrf <- ggplot(or_ndc_county_time_series_week, aes(y =wrf_smk_pm, x = date)) +
           geom_jitter(color="red")
# print(plot_smoke_ts_wrf)

plot_smoke_ts_krig <- ggplot(or_ndc_county_time_series_week, aes(y =krig_smk_pm, x = date)) +
           geom_jitter() 
# print(plot_smoke_ts_krig)

plot_smoke_ts_gwr <- ggplot(or_ndc_county_time_series_week, aes(y =geo_smk_pm, x = date)) +
           geom_jitter() 
# print(plot_smoke_ts_gwr)

plot_smk <- ggplot(or_ndc_county_time_series_week, aes(x = date)) +
  geom_jitter(aes(y =wrf_smk_pm, x = date), color='green', size = 1) + 
  geom_jitter(aes(y =krig_smk_pm, x = date), color='blue', size = 1) + 
  geom_jitter(aes(y =geo_smk_pm, x = date), color="red", size = 1)  +
  ylab("Smoke PM 2.5")

print(plot_smk)

### County
or_smk <- or_ndc_county_time_series_week %>% 
  select(county, date, wrf_smk_pm, krig_smk_pm, geo_smk_pm) %>% 
  gather(key = smk_method, value = pm, -date, -county)

or_smk_plot <- ggplot(or_smk, aes(x = date, y = pm)) +
  geom_point(aes(colour = smk_method), alpha = 0.8, size = 0.8) +
  facet_wrap(~county) +
  ggtitle("Oregon: 2013 Fire Season Smoke PM2.5") +
  ylab("Smoke PM2.5 ug/m^3") +
  xlab("Year: 2013") +
  theme_bw()

print(or_smk_plot)

ggsave("../plot_new/ndc_smoke_county.pdf", plot = plot_mult_week, 
       width = 12, height = 8, units = "in")

```

## Distributed Lag Model

### Lag 1

```{r DLM, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
asthma_ndc_df <- data.frame(asthma_ndc)

asthma_ndc_df <- asthma_ndc_df %>% # GWR
  filter(!is.na(geo_wt_pm_lag7_zip)) # 503998

which(colnames(asthma_ndc_df)=="geo_wt_pm_zip") # 88
which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") #115

asthma_ndc_lag <- asthma_ndc_df %>%
  select(88, 115:121) # GWR-method

lag_max <- 7
n <- length(asthma_ndc_lag[,1])

asthma_ndc_lag <- as.matrix(asthma_ndc_lag)

B <- ns(0:lag_max, df =4, intercept=TRUE)

# the nrow B should equal ncol(pm)
dim(B)
dim(asthma_ndc_lag)

asthma_ndc_lag_B <- asthma_ndc_lag%*%B
fit <- clogit(outcome ~ asthma_ndc_lag_B + wrf_temp_zip + strata(personkey), asthma_ndc_df)
coef(fit)
summary(fit)

dlparms <- grep("asthma_ndc_lag_B",names(coef(fit)))
DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96

DLestimate$estimate <- exp(DLestimate$estimate)
DLestimate$lower <- exp(DLestimate$lower)
DLestimate$upper <- exp(DLestimate$upper)

DLestimate

DLestimate$lag <- 0:lag_max
p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
p <- p + geom_line(aes(x=lag), linetype=2, size=2)
p <- p + theme_bw()
p

DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])

cumulative <- sum(DLestimate$estimate)
cumulative
## [1] 0.009141771
cumulative_se <- sqrt(sum(DLvar))
cumulative_se
## [1] 0.000955469
cumulative_CI <- cumulative + cumulative_se * c(-1.96, 1.96)
cumulative_CI
## [1] 0.007269052 0.011014490

```

```{r overall lag 1, warning =F, echo = F, results='asis', eval=FALSE} 
# dataframe list

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
# asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "Asthma Inhalers Lag 1"
outcome_name <- "Asthma Inhalers Lag 1"

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
which(colnames(asthma_ndc_df)=="wrf_smk_pm_lag1_zip") # 108
which(colnames(asthma_ndc_df)=="krig_pm_lag1_zip") # 122
which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") # code to find column numbers, 115
pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases

# new loop for age categories

# empty matrix (12 x 10 matrix)
point_estimates <- matrix(nrow = 3, ncol = 9, byrow = T)

colnames(point_estimates) <- c('outcome', 'pm_method', 'n', 'n_events', 
                               'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

# fill in the outcome namedataframe before method loop
point_estimates[, 1] <- outcome_name

# dataframe for analysis creation
# bind columns back together 
df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
  # remove missing pm values
  filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
  filter(!is.na(krig_pm_lag1_zip)) %>% 
  filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
  # the following code makes sure that the counterfactual values retained are 
  # symetric in that number of obs before = number of obs after
  mutate(obs_diff_admission = (fromdate - date)/7) 
# dataframe is already for the entire fire season, so I don't need to subset anymore

# second loop to run a model for each pm estimation method
for(j in 38:40){

  # variable to model 
  var_name <- colnames(df_analysis[j])

  # set row number to fill
  row_n <- j-37

  # only run the model if the dataframe has observations
  if(nrow(df_analysis) != 0){
    # conditional logistic regression model
    mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

    # populate matrix
    point_estimates[row_n, 2] <- method_list[row_n]
    point_estimates[row_n, 3] <- mod$n
    point_estimates[row_n, 4] <- mod$nevent
    # odds ratio
    point_estimates[row_n, 5] <- round(exp(summary(mod)$coefficient[1,1]), 3)

    # 95% lower bound
    point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # 95% upper bound
    point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) +
                                             1.96*(summary(mod)$coefficient[1,3])), 3)
    # standard error
    point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,3], 4)
    # p val
    point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,5], 4)

    # create else statement that fills matrix with missing so I still have the row
    # in the final dataframe
  } else {point_estimates[row_n, 2] <- method_list[row_n]
  point_estimates[row_n, 3] <- 0
  point_estimates[row_n, 4] <- 0
  point_estimates[row_n, c(5:9)] <- 99 } # end 'if else' statement
} # end methods loop

# save point estimates as a dataframe
combined_point_est_df <- as_data_frame(point_estimates)

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" |
  #      pm_method == "WRF-Chem Smoke") %>% 
  # subset columns I want to put in to the table
  select(2, 4:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 three smoke method and pharmacy outcomes",
                 # row group by outcome
                 rgroup = "Asthma Inhalers Lag 1",
                 n.rgroup = c(rep(3, 1)), # 6 rows for each age cat for each outcome
                 # column headers
                 header = c("Est Method", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(3, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llcccc", # column alignment,
                 tfoot="&dagger; Referent periods matched to events on same day of week within May to September fire season."
) # end table

print(tab)

combined_point_est_df_new <- combined_point_est_df %>%
  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
                               ifelse(as.numeric(n_events)<100, "0", NA))) %>%
  select(-n_events) %>%
  rename(n_events = n_events_new)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_age_plot <- combined_point_est_df_new %>% 
  # filter columns to just geo-smk
  #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                          99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                          99, upper95)) %>% 
  # subset columns I want to put in to the table
  select(1, 2, 9, 4:6) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
                                   levels = unique(wrf_geo_age_plot$outcome))

wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
                                     levels = unique(wrf_geo_age_plot$pm_method))

wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)

## ggplot 
print_plot <- ggplot(wrf_geo_age_plot,
                     aes(x = pm_method, y = odds_ratio, color = pm_method), na.rm = T) +
  geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red",  "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~outcome, nrow = 3, scales = "free") +
  geom_hline(yintercept = 1, linetype=2) +
  #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
  #ylim(0, 2) +
  xlab('Smoke Estimation Method') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        # axis element
        #axis.text.x = element_blank(),
        #axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/ndc_overall_lag1_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r time stratified age strata lag 1, warning =F, echo = F, results='asis', eval=FALSE} 
# dataframe list

method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "asthma inhalers Lag 1"
outcome_name <- "Asthma inhalers Lag 1"

# age category list
age_cat_list <- c(0,1,2)

# create an empty list to row bind dataframes together
datalist1 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases

# new loop for age categories
for(k in 0:2){
  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

  colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
                                 'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome namedataframe before method loop
  point_estimates[, 1] <- outcome_name
  # repeat the age category 4 times for each pm method
  point_estimates[, 2] <- k

  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
    # limit to specific age category
    filter(age_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
  for(j in 38:40){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # set row number to fill
    row_n <- j-37

    # only run the model if the dataframe has observations
    if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -                                                                             1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                               1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
    } else {point_estimates[row_n, 3] <- method_list[row_n]
    point_estimates[row_n, 4] <- 0
    point_estimates[row_n, 5] <- 0
    point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
  } # end methods loop

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)
  datalist1[[k+1]] <- point_est_df
} # end age category loop

# bind rows of age category estimates together
combined_point_est_df <- bind_rows(datalist1)

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" |
  #      pm_method == "WRF-Chem Smoke") %>% 
  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
                           ifelse(age_cat == 1, "15-65",
                                  ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
                 # row group by outcome
                 rgroup = "Asthma inhalers Lag 1",
                 n.rgroup = c(rep(9, 1)), # 6 rows for each age cat for each outcome
                 # column headers
                 header = c("Est Method", "Age Group", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llcccc", # column alignment,
                 tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within May to September fire season."
) # end table

print(tab)

combined_point_est_df_new <- combined_point_est_df %>%
  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
                               ifelse(as.numeric(n_events)<100, "0", NA))) %>%
  select(-n_events) %>%
  rename(n_events = n_events_new)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_age_plot <- combined_point_est_df_new %>% 
  # filter columns to just geo-smk
  #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                          99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                          99, upper95),
         age_cat2 = ifelse(age_cat == 0, "<15", 
                           ifelse(age_cat == 1, "15-65",
                                  ifelse(age_cat == 2, ">65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(1, 3, 11, 10, 5:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
                                   levels = unique(wrf_geo_age_plot$outcome))

wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
                                     levels = unique(wrf_geo_age_plot$pm_method))

wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
                                    levels = unique(wrf_geo_age_plot$age_cat2))

wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)

## ggplot 
print_plot <- ggplot(wrf_geo_age_plot,
                     aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
  geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red",  "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~outcome, nrow = 3, scales = "free") +
  geom_hline(yintercept = 1, linetype=2) +
  #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
  #ylim(0, 2) +
  xlab('Age Category ') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10), 
        # axis element
        #axis.text.x = element_blank(),
        #axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/ndc_age_fig_three_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```

```{r sex strata cdc met  lag 1, warning = F, echo = F, results='asis', eval=FALSE} 
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

# outcome name
outcome <- "asthma inhalers Lag 1"
outcome_name <- "Asthma inhalers Lag 1"

# create an empty list to row bind dataframes together
datalist1 <- list()

# sex category list
sex_strata_list <- c(0,1)

# data wrangling ----
# Producing conditional logit model estimates loop

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- asthma_ndc_df[, c(108, 122, 115, 93)]/10  # create 10 unit increases

# new loop for sex categories
for(k in 0:1){

  # empty matrix (12 x 10 matrix)
  point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

  colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
                                 'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

  # fill in the outcome namedataframe before method loop
  point_estimates[, 1] <- outcome_name
  # repeat the sex category 4 times for each pm method
  point_estimates[, 2] <- k

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    # remove missing pm values
    filter(!is.na(wrf_smk_pm_lag1_zip)) %>% 
    filter(!is.na(krig_pm_lag1_zip)) %>% 
    filter(!is.na(geo_wt_pm_lag1_zip)) %>% 
    # limit to specific age category
    filter(sex_ind == k) %>% 
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (fromdate - date)/7) 
  # dataframe is already for the entire fire season, so I don't need to subset anymore

  # second loop to run a model for each pm estimation method
  for(j in 38:40){

    # variable to model 
    var_name <- colnames(df_analysis[j])

    # set row number to fill
    row_n <- j-37

    # only run the model if the dataframe has observations
    if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                               1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                               1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
    } else {point_estimates[row_n, 3] <- method_list[row_n]
    point_estimates[row_n, 4] <- 0
    point_estimates[row_n, 5] <- 0
    point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement
  } # end methods loop

  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)

  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist1[[k+1]] <- point_est_df
} # end sex category loop

# bind rows of sex category estimates together
combined_point_est_df <- bind_rows(datalist1)

wrf_geo_sex <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
                          ifelse(sex == 1, "Female", NA))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8)%>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
                 caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and pharmacy outcomes stratified by sex",
                 # row group by outcome
                 rgroup = "Asthma inhalers Lag 1",
                 n.rgroup = c(rep(6, 1)), # 2 rows for each sex strata for each outcome
                 # column headers
                 header = c("Est Method", "Sex Strata", "Events",
                            "OR&dagger;", "Lower", "Upper"),
                 # column spanner
                 cgroup = c("", "95% CI"), 
                 n.cgroup = c(4, 2),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llcccc", # column alignment,
                 tfoot="&dagger; Time-stratified: referent periods matched to events on same day of week within July to October fire season."
) # end table

print(tab)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_sex_plot <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  #  filter(pm_method == "GWR Smoke" | 
  #       pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
                          ifelse(sex == 1, "Female", NA))) %>%
  # subset columns I want to put in to the table
  select(1, 3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
                                   levels = unique(wrf_geo_sex_plot$outcome))

wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
                                     levels = unique(wrf_geo_sex_plot$pm_method))

wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
                                   levels = unique(wrf_geo_sex_plot$sex_cat))

wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)

## ggplot
print_plot <- ggplot(wrf_geo_sex_plot,
                     aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
  geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
  # custom color 
  scale_color_manual(name = "Smoke-Estimation Method", 
                     values = c("red", "blue", "#32115C"),
                     guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~outcome, nrow = 3, scales = "free") +
  geom_hline(yintercept = 1, linetype=2) +
  #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
  ylab(expression(paste("Odds Ratio for 10Âµg/m"^3, " Increase in PM"[2.5]))) +
  #ylim(0, 2) +
  xlab('Sex') +
  # plot theme
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        # axis element
        #axis.text.x = element_blank(),
        #axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(angle = 90),
        # legend elements
        legend.position = "bottom")
#legend.text = element_text(size = 8))

print(print_plot)
# save figure
ggsave("../plot_lag/ndc_sex_fig_three_lag1.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

```