### 2017-11-06 r_scripts/analysis/model_dl_spline_knots.Rmd

---
title: "Oregon distributed lag model best fit degree of freedom"
author: "Jingyang Liu"
date: "2017/11/6"
output: html_document
---

```{r library, include=FALSE, echo=FALSE}
library(tidyverse)
# library(data.table)
library(survival)
library(ggplot2)
library(htmlTable)
library(knitr)
library(broom)
library(lme4)
library(maps)
library(lubridate) # date
library(ggrepel)
library(sp)
library(cowplot)
library(geofacet)

# distributed lag model libraries
library(dlnm)
library(mvmeta)
library(splines)

```

Load data

```{r descriptive table, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip.csv")
inhaler_ndc <- read_csv(read_path6)
# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis

# dataframe to loop through
inhaler_ndc_df <- data.frame(inhaler_ndc)

```

```{r input, echo = FALSE, message=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
or_pop <- read_csv(read_path9)

or_pop <- or_pop %>%
  rename(county = GEO_display_label, pop = respop72013) %>%
  select(county, pop) 
or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))

inhaler_ndc_lag_df <- inhaler_ndc_df %>%
  left_join(or_zip_to_county, by = "ZIPCODE") %>%
  left_join(or_pop, by = "county") %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA))) %>%
  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)

inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"

```

### Find the smallest of AIC with df from 3 to 6

Choose df = 5 because it has the least AIC.

```{r, message=FALSE, echo = FALSE, warning=FALSE}

# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
# which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160
# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_zip") # 93
# which(colnames(inhaler_ndc_lag_df)=="wrf_temp_lag1_zip") # 150

inhaler_geo_lag <- inhaler_ndc_lag_df %>%
  select(159, 160:166) # GWR-method

inhaler_temp_lag <- inhaler_ndc_lag_df %>%
  select(93, 150:156) # GWR-method

inhaler_geo_lag <- as.matrix(inhaler_geo_lag)
inhaler_temp_lag <- as.matrix(inhaler_temp_lag)

# create fit matrix
fit_mat <- matrix(NA, nrow=4, ncol=5)
colnames(fit_mat) <- c("outcome", "exposure", "fit", "df", "aic")

# fill outcome name in first column
fit_mat[,1] <- "Beta - 2 agonists"
# fill in exposure (geo smoke)
fit_mat[,2] <- "geo_smk10"
# fit type
fit_mat[,3] <- "ns"

# 2 df will not work for a spline, going from 3 to 6 df
  for(h in 3:6){ # start loop for df

    # fitting a natural spline for lag with 5 knots?
    fit_mat[h-2, 4] <- degree_freedom <- h

    # define basis b using natural spline function
    B <- ns(0:7, df = degree_freedom, intercept = T)

    # multiply lagged pm matrix by basis
    inhaler_geo_lag_B <- inhaler_geo_lag%*%B
    inhaler_temp_lag_B <- inhaler_temp_lag%*%B

    fit <- clogit(outcome ~ inhaler_geo_lag_B + inhaler_temp_lag_B + strata(personkey), 
              inhaler_ndc_lag_df)

    AIC(fit)

    # fill AIC
    fit_mat[h-2,5] <- round(AIC(fit),5)

    } # end model fit loop

# now that we have a range of df/knots, I want to filter to the minimum aic
# for each outcome and print out that value in a table

kable(fit_mat, caption="Distributed lag spline of 3-6 degree of freedom fit by AIC")

lag_spline_best_fit <- as_tibble(fit_mat) %>%  
  slice(which.min(aic)) # df of 3 has the smallest AIC

# kable
knitr::kable(lag_spline_best_fit, caption = paste0("Distributed lag spline ",
  "degree of freedom best fit by AIC"))
```