### 2017-11-20 r_scripts/analysis/oregon_cvd_dl_over65_results.Rmd
started analysis looking at cvd dl outcomes in the over 65 group.

---
title: Distributed lag effects of smoke on cardiovascular disease in persons over 65
author: "Ryan Gan"
date: "2017-11-20"
output: html_document
---