### 2017-12-05 r_scripts/analysis/agu_saba_timeseries.Rmd
Made figure for jeff and kate.

---
title: "AGU Timeseries"
author: "Ryan Gan"
date: "12/1/2017"
output: html_document
---

```{r setup, echo=F, warning=F, message=F}
library(tidyverse) # general data wrangle

# knitr options
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

## Document Purpose

Map for Kate and Jeff for AGU to demonstrate variability in health outcomes.

```{r data wrangle}
# saba timestratifed to counts for fills with a icd9 as well
saba <- read_csv("../../data/health/2013-oregon_casecross_saba.csv") %>% 
  # get rid of referent values
  filter(outcome == 1)

# time sereis of saba
saba_ts <- saba %>% 
  group_by(MSA, date) %>% 
  summarise(saba_n = n()) %>% 
  mutate(metroarea = case_when(
    MSA == 13460 ~ "Bend-Redmond",
    MSA == 18700 ~ "Corvallis",
    MSA == 21660 ~ "Eugene-Springfield",
    MSA == 32780 ~ "Medford",
    MSA == 38900 ~ "Portland-Vancouver-Beaverton",
    MSA == 41420 ~ "Salem",
    MSA == 41999 ~ "Not in MSA"),
    day = weekdays(date),
    weekend = ifelse(day %in% c("Saturday", "Sunday"), "Weekend", "Weekday"),
    week_n = lubridate::floor_date(date, unit = "week"))

# calulate median weekend and weekday value
saba_center_stats <- saba_ts %>% 
  # group by MSA and weekend
  group_by(metroarea, weekend) %>% 
  summarise(saba_median = median(saba_n), saba_mean = mean(saba_n),
            saba_sd = sd(saba_n))

# read county pm time sereis
# pm path
pm_path <- paste0("../../data/pm/2013-oregon_county_pm25.csv")
# county pm
county_pm_df <- read_csv(pm_path) %>% 
  mutate(county = stringr::str_replace(county, "[.]", ""))
```

Plot of SABA counts by metro areas in Oregon for 2013 wildfire season (May through September). I couldn't easily get counts in a county out of the data as we had to code county based on reported ZIP code and we never added this as a covariate to the dataset (I never explicitly asked Jingyang to do this; I plan to, but it will probably take a bit).

But, I think MSA works pretty well to illustrate variability in the SABA health outcome that may be linked with smoke.

### Figure 1: Time-series of short-acting beta-2 agonists over the metropolitan statistical areas for the state of Oregon.

```{r msa plot}
plot <- ggplot(saba_ts, 
    aes(x=date, y=saba_n, group = weekend, color = weekend)) +
  geom_point() +
  geom_hline(data=saba_center_stats, aes(yintercept=saba_mean, group=weekend,
    color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  facet_wrap(~metroarea) +
  ylab("SABA Count") +
  xlab("Date") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_line(colour="grey"),
    #panel.grid.minor = element_blank(),
    legend.position = "bottom",
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"))

plot
```

I've produced plots of SABA counts by MSA area. If you look at the major MSA areas, you'll see that in the larger MSAs like Portland, there is a periodic seasonal increase in spring SABA counts that fall off in summer and increase again around fall. This is the same type of trend seen for asthma. Some of the other more populated counties begin to exhibit this shape too, but not as pronounced as the more populated areas. If you look at Medford (the county impacted to the greatest extent by the Douglas-Complex fire, there is a little increase around the time of the fire (started: July 26th, contained: September 3rd 2013). If you look at some of the other MSA trends, they do not have this little peak around this time period.

Note I had a MSA code that I was not able to link with an MSA so I'm guessing it was not in a MSA; data key was not explicit about this but I'm pretty sure it's for fills for people not in MSA. Also note that there are some weekdays that have similar fill-counts as weekends; these are Memorial Day, July 4th, and Labor Day holidays.

Limiting to MSA timeseries to Medford MSA, which is in Jackson county.

### Figure 2: Time-series of SABA limited to Medford MSA.

```{r medford}
medford_center_stats <- saba_center_stats %>% 
  filter(metroarea == "Medford")

medford_ts <- saba_ts %>% 
  filter(metroarea == "Medford") %>% 
  full_join(medford_center_stats, by = c("metroarea", "weekend")) %>% 
  mutate(diff_mean = saba_n - saba_mean,
    saba_z = (saba_n-saba_mean)/saba_sd, 
    diff_med = saba_n - saba_median,
    fire_period = ifelse(date >= "2013-07-26" & date <= "2013-09-03",50,NA))
```

Medford MSA SABA count time-series plot for Weekend and Workweek fills. Dotted verticle lines represent the start date of the Douglas-Complex fires and the date it was considered 100% contained.

### Figure 3: Time-series of SABA Z-Score.

```{r medford count timeseries}
# medford time series
ts_plot <- ggplot(medford_ts, 
  aes(x=date, y=saba_n, group=weekend, color=weekend)) +
  geom_point() +
  geom_hline(data=medford_center_stats, aes(yintercept=saba_mean, color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylim(0, 50) +
  ylab("SABA Count") +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom")

ts_plot
```

Plotting the Z-score of SABA counts (observed - mean)/standard deviation. Z-score accounted for weekend/weekday difference

```{r saba z}
# z-score plot
z_plot <-ggplot(medford_ts, 
  aes(x=date, y=saba_z, group=weekend, color=weekend)) +
  geom_point() +
  geom_hline(yintercept=0, color="red") +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylim(-3, 3) +
  ylab("SABA Z-Score") +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme_minimal()
z_plot
```

Limiting population-weighted PM~2.5~ time-series to Jackson county. Jackson county population-weighted geographically-weighted ridge regression PM~2.5~ time-series.

### Figure 4: Truncated SABA count

I'm truncating to July to October.

```{r trunc saba ts}
jul_oct_medford <- medford_ts %>% 
  filter(date >= "2013-07-01" & date <= "2013-10-01")

jul_oct_central_vals <- medford_ts %>% 
  group_by(weekend) %>% 
  summarise(saba_med = median(saba_n))

# medford time series
saba_plot <- ggplot(jul_oct_medford, 
  aes(x=date, y=saba_n, group=weekend, color=weekend)) +
  geom_point() +
  #geom_hline(data=jul_oct_central_vals , aes(yintercept=saba_med, color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylim(0, 40) +
  ylab("SABA Count") +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"))

saba_plot
```

### Figure 5: Time-series of GWR PM~2.5~ values.

```{r gwr smk timeseries}
# jackson plot
jackson_pm <- county_pm_df %>% 
  filter(county == "Jackson") %>% 
  mutate(smk10 = ifelse(geo_wt_pm > 10, 1, 0),
    fire_period = ifelse(date >= "2013-07-26" & date <= "2013-09-03",1,0)) %>% 
  filter(date >= "2013-07-01" & date <= "2013-10-01")

# plot 
pm_plot <- ggplot(jackson_pm, aes(x=date, y=geo_wt_pm)) +
  geom_point(color = "#000046") +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylab(expression(paste("PM"[2.5], "  ", mu, "g/", m^3, sep=""))) +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

pm_plot  
```

### Figure 6: Binding truncated SABA counts and PM~2.5~

```{r trunc plot}
cowplot::plot_grid(pm_plot, saba_plot,
  #labels = c("PM2.5 Exposure", "Rescue Inhaler"), 
  nrow = 2, align = "v")
```