### 2017-12-07 r_scripts/analysis/asthma_cohort_descriptive_epi.Rmd
working on descriptives for asthma cohort

---
title: "Asthma Cohort Descriptive Epidemiology"
author: "Ryan Gan"
date: "2017-12-06"
output: html_document
---

## Overview

Descriptive statistics of the assembled asthma cohort. This cohort is made up of all billing records for procedures or pharamacy fills for anyone in the Oregon APAC data that had been billed for a diagnosis of asthma at any visit in 2013.

```{r setup, echo =F, warning=F, message=F}
library(tidyverse) # general data wrangle

# load outcomes vectors
load("./data/health/outcome_list.RData")
# subset asthma and saba vector
asthma_icd9 <- pluck(outcome_icd9_list, "asthma")
saba_ndc <- pluck(outcome_icd9_list, "saba")

# knitr options
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

Read cohort dataset (reading in a small chunk for now.)

```{r cohort_import}
# read path
read_path <- "./data/health/2013-oregon_asthma_cohort.csv"
# read in asthma cohort (reading a million rows for now)
asthma_cohort <- data.table::fread(read_path, 
  # read options
  header = T, nrows = 10000, colClasses = rep("character", 72)) %>% 
  # replace *NULL* with NA
  mutate_all(funs(replace(., .== "*NULL*" | . == "", NA))) %>% 
  # create visit type indicator 
  mutate(visit_type = as.factor(
    case_when((dx1 %in% asthma_icd9) ~ "dx_asthma_primary",
    # asthma but not as a primary
    !(dx1 %in% asthma_icd9) & (dx2 %in% asthma_icd9 | dx3 %in% asthma_icd9 | 
      dx4 %in% asthma_icd9 | dx5 %in% asthma_icd9 | dx6 %in% asthma_icd9 | 
      dx7 %in% asthma_icd9 | dx8 %in% asthma_icd9 | dx9 %in% asthma_icd9 | 
      dx10 %in% asthma_icd9 | dx11 %in% asthma_icd9 | dx12 %in% asthma_icd9 | 
      dx13 %in% asthma_icd9) ~ "dx_asthma_not_primary",
    # other dx 
    !(dx1 %in% asthma_icd9 | dx2 %in% asthma_icd9 | dx3 %in% asthma_icd9 | 
      dx4 %in% asthma_icd9 | dx5 %in% asthma_icd9 | dx6 %in% asthma_icd9 | 
      dx7 %in% asthma_icd9 | dx8 %in% asthma_icd9 | dx9 %in% asthma_icd9 | 
      dx10 %in% asthma_icd9 | dx11 %in% asthma_icd9 | dx12 %in% asthma_icd9 | 
      dx13 %in% asthma_icd9) & is.na(ndc) ~ "dx_other",
    # ndc code
    (ndc %in% saba_ndc) ~ "pharm_saba",
    !is.na(ndc) & !(ndc %in% saba_ndc) ~ "pharm_other")))
```