### 2018-01-09 r_scripts/analysis/asthma_care_utilization.Rmd
added new file.

---
title: "Multiple visits in persons likely to have asthma"
author: "Ryan Gan"
date: "1/9/2018"
output: html_document
---

## Document Purpose

This document attempts to describe utilization of healthcare in persons who are likely to have asthma. It will eventually see if wildfire smoke may influence utilization.

First step is to read in the persons with asthma in the fireseason. I may end up doing this for the entire year, but that dataset is much larger and would take a bit.

```{r setup, echo = F, warning = F, message = F}
# libraries.
library(tidyverse)

# knitr options
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```