### 2018-06-07 r_scripts/analysis/msa_asthma_timeseries.Rmd
updated msa smoke and asthma care plot.

---
title: "MSA Asthma Place of Service Timeseries"
author: "Ryan Gan"
date: "6/7/2018"
output: html_document
---

Libraries.

```{r library}
# libraries
library(tidyverse) # for ggplot and general data wrangle
```

Loading case-crossover dataframe list and SABA fill list.

```{r load_data}
# load asthma casecross place of service list
load("../../data/health/asthma_cc_pos_list.RData")
# extract names
vars_keep <- names(asthma_cc_pos_list[[4]])
```

Load SABA DF.

```{r saba_df}
saba = read_csv("../../data/health/2013-oregon_casecross_saba.csv") %>% 
  rename(identifier = personkey, ZIP = ZIPCODE) %>% 
  mutate(service_place = "SABA Fill",
         dx1 = ndc) %>% 
  # keeping only vars in the other casecrossover dataframes
  dplyr::select(vars_keep, ndc) %>% 
  # turn in to list
  list()

# assign name
names(saba) = "saba"
```

Load Jingyang's mothly count to extract populations for MSA.
```{r msa_pop}
# read saba and asthma ed count# read  
population <- read_csv("./data/health/saba_month_counts.csv") %>% 
  select(msa_name, POPESTIMATE2013) %>% 
  rename(metroarea = msa_name,
         pop = POPESTIMATE2013) %>% 
  unique()
```

Subset to only outcomes and add metroarea MSA.

```{r list_subset}
# adding in age category to each dataframe in list 
# (i should have done this earlier)
asthma_pos_count <- c(asthma_cc_pos_list, saba) %>% 
  map_dfr(., function(df){
    return_df <- df %>% 
      # remove unknown gender
      filter(outcome ==1 ) %>% 
      mutate(metroarea = case_when(MSA == 13460 ~ "Bend",
                                MSA == 18700 ~ "Corvallis",
                                MSA == 21660 ~ "Eugene",
                                MSA == 32780 ~ "Medford",
                                MSA == 38900 ~ "Portland",
                                MSA == 41420 ~ "Salem",
                                MSA == 41999 ~ "Not in MSA"),
        service_place = case_when(service_place == "Emergency Room – Hospital" ~ 
                                    "Emergency Department", 
          TRUE ~ service_place)) %>% 
      # group by date, metroarea, and service place
      group_by(date, metroarea, service_place) %>% 
      summarise(n_events = sum(outcome))
  })
```

```{r medford_plot}
check_ts <- asthma_pos_count %>% 
  filter(metroarea == 'Medford')

ggplot(check_ts, aes(x=date, y=n_events, group = metroarea, color=metroarea)) +
  geom_point() +
  facet_wrap(~service_place, scales = "free_y")
```

Weekly sum.

```{r weekly_count}
asthma_weekly_count <- asthma_pos_count %>% 
  mutate(week_date = lubridate::ceiling_date(date, unit = 'week')) %>% 
  group_by(week_date, metroarea, service_place) %>% 
  summarise(n_events = sum(n_events)) %>% 
  left_join(population, by = 'metroarea') %>% 
  mutate(rate_100k = (n_events/pop)*100000) %>% 
  filter(service_place != 'Ambulance-Land') %>% 
  filter(week_date != '2013-10-06') %>% 
  filter(!is.na(pop))
```

Weekly count plot.I actually like this plot more I think.

```{r pos_plot}
pos_plot = ggplot(asthma_weekly_count, 
  aes(x=week_date, y=rate_100k, group = service_place, color=service_place)) +
  geom_point(size = .7) +
  scale_color_manual('Care Utilization', 
    values = c("#3c1053", "#7f00ff", "#e100ff", "#00b09b", "#0ed2f7", "#c31432")) +
  facet_wrap(~metroarea, ncol=1) +
    ylab("Weekly Rate per 100,000 Persons") +
    xlab("Date") +
    ggtitle('B. Asthma Care Utilization') +
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        strip.text = elemnt_text(vjust= -1),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_text(angle = 90))

pos_plot
```

Read PM2.5 and create MSA PM

```{r}
# read saba
path = paste0('../../data/health/or_respiratory_may_to_sep_time_strat_casecross_er.csv')

# read respiratory casecross and find unique zipcodes by MSA values
zip_msa = read_csv(path, col_types = cols(.default = "c")) %>% 
    select(ZIPCODE, MSA) %>% 
    unique()

# read zip pm; join with msa
msa_pm = read_csv('./data//pm/2013-oregon_zip_pm25.csv', 
                  col_types = cols(ZIPCODE = 'c')) %>% 
    # join with unique MSA vector
    left_join(zip_msa, by = 'ZIPCODE') %>%
    # mutate
    mutate(ZIPCODE = as.factor(ZIPCODE),
        MSA = as.factor(MSA),
        # assign metro name to number
        metroarea = case_when(MSA == 13460 ~ "Bend",
                              MSA == 18700 ~ "Corvallis",
                              MSA == 21660 ~ "Eugene",
                              MSA == 32780 ~ "Medford",
                              MSA == 38900 ~ "Portland",
                              MSA == 41420 ~ "Salem"))  %>% 
    # filter to zips in an MSA only
    filter(!is.na(metroarea)) 
```

Plot of PM2.5 by MSA.

```{r pm_plot}
# plot of pm by msa
pm_msa_plot = ggplot(data = msa_pm, aes(x = date, y = geo_smk_pm, group = ZIPCODE)) +
    geom_line(color = '#e94057') +
    # facet by metro area
    facet_wrap(~metroarea, ncol = 1) +
    # y label
    ylab(expression(paste("Wildfire Smoke PM"[2.5]," 10µg/m"^3))) +
    # x label
    xlab("Date") +
    ggtitle(expression(paste('A. Wildfire Smoke PM'[2.5]))) +
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_text(angle = 90))

pm_msa_plot
```