### 2018-06-11 r_scripts/analysis/msa_asthma_timeseries.Rmd
updated some stuff.

---
title: "MSA Asthma Place of Service Timeseries"
author: "Ryan Gan"
date: "6/7/2018"
output: html_document
---

Libraries.

```{r library}
# libraries
library(tidyverse) # for ggplot and general data wrangle
```

Loading case-crossover dataframe list and SABA fill list.

```{r load_data}
# load asthma casecross place of service list
load("../../data/health/asthma_cc_pos_list.RData")
# extract names
vars_keep <- names(asthma_cc_pos_list[[4]])
```

Load SABA DF.

```{r saba_df}
saba = read_csv("../../data/health/2013-oregon_casecross_saba.csv") %>% 
  rename(identifier = personkey, ZIP = ZIPCODE) %>% 
  mutate(service_place = "SABA Fill",
         dx1 = ndc) %>% 
  # keeping only vars in the other casecrossover dataframes
  dplyr::select(vars_keep, ndc) %>% 
  # turn in to list
  list()

# assign name
names(saba) = "saba"
```

Load Jingyang's mothly count to extract populations for MSA.
```{r msa_pop}
# read saba and asthma ed count# read  
population <- read_csv("./data/health/saba_month_counts.csv") %>% 
  select(msa_name, POPESTIMATE2013) %>% 
  rename(metroarea = msa_name,
         pop = POPESTIMATE2013) %>% 
  unique()
```

Subset to only outcomes and add metroarea MSA.

```{r list_subset}
# adding in age category to each dataframe in list 
# (i should have done this earlier)
asthma_pos_count <- c(asthma_cc_pos_list, saba) %>% 
  map_dfr(., function(df){
    return_df <- df %>% 
      # remove unknown gender
      filter(outcome ==1 ) %>% 
      mutate(metroarea = case_when(MSA == 13460 ~ "Bend",
                                MSA == 18700 ~ "Corvallis",
                                MSA == 21660 ~ "Eugene",
                                MSA == 32780 ~ "Medford",
                                MSA == 38900 ~ "Portland",
                                MSA == 41420 ~ "Salem",
                                MSA == 41999 ~ "Not in MSA"),
        service_place = case_when(service_place == "Emergency Room – Hospital" ~ 
                                    "Emergency Department", 
          TRUE ~ service_place)) %>% 
      # group by date, metroarea, and service place
      group_by(date, metroarea, service_place) %>% 
      summarise(n_events = sum(outcome))
  })
```

```{r medford_plot}
check_ts <- asthma_pos_count %>% 
  filter(metroarea == 'Medford')

ggplot(check_ts, aes(x=date, y=n_events, group = metroarea, color=metroarea)) +
  geom_point() +
  facet_wrap(~service_place, scales = "free_y")
```

Weekly sum.

```{r weekly_count}
asthma_weekly_count <- asthma_pos_count %>% 
  mutate(week_date = lubridate::ceiling_date(date, unit = 'week')) %>% 
  group_by(week_date, metroarea, service_place) %>% 
  summarise(n_events = sum(n_events)) %>% 
  left_join(population, by = 'metroarea') %>% 
  mutate(rate_100k = (n_events/pop)*100000) %>% 
  filter(service_place != 'Ambulance-Land') %>% 
  filter(week_date != '2013-10-06') %>% 
  filter(!is.na(pop))
```

Weekly count plot.I actually like this plot more I think.

```{r pos_plot}
pos_plot <- ggplot(asthma_weekly_count, 
  aes(x=week_date, y=rate_100k, group = service_place, color=service_place)) +
  geom_point(size = 1.25) +
  scale_color_manual('Care Utilization', 
    values = c("#3c1053", "#7f00ff", "#e100ff", "#00b09b", "#0ed2f7", "#c31432")) +
  facet_wrap(~metroarea, ncol=1) +
    ylab("Weekly Rate per 100,000 Persons") +
    xlab("Date") +
    ggtitle('B. Asthma Care Utilization') +
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
                                        size = 0.5),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_text(angle = 90),
        legend.key = element_blank())

pos_plot
```

Read PM2.5 and create MSA PM

```{r msa_pm}
# read saba
path = paste0('../../data/health/or_respiratory_may_to_sep_time_strat_casecross_er.csv')

# read respiratory casecross and find unique zipcodes by MSA values
zip_msa = read_csv(path, col_types = cols(.default = "c")) %>% 
    select(ZIPCODE, MSA) %>% 
    unique()

# read zip pm; join with msa
msa_pm <- read_csv('./data//pm/2013-oregon_zip_pm25.csv', 
                  col_types = cols(ZIPCODE = 'c')) %>% 
    # join with unique MSA vector
    left_join(zip_msa, by = 'ZIPCODE') %>%
    # mutate
    mutate(ZIPCODE = as.factor(ZIPCODE),
        MSA = as.factor(MSA),
        # assign metro name to number
        metroarea = case_when(MSA == 13460 ~ "Bend",
                              MSA == 18700 ~ "Corvallis",
                              MSA == 21660 ~ "Eugene",
                              MSA == 32780 ~ "Medford",
                              MSA == 38900 ~ "Portland",
                              MSA == 41420 ~ "Salem"))  %>% 
    # filter to zips in an MSA only
    filter(!is.na(metroarea)) 
```

Plot of PM2.5 by MSA.

```{r pm_plot}
# plot of pm by msa
pm_msa_plot <- ggplot(data = msa_pm, aes(x = date, y = geo_smk_pm, group = ZIPCODE)) +
    geom_line(color = '#e94057') +
    # facet by metro area
    facet_wrap(~metroarea, ncol = 1) +
    # y label
    ylab(expression(paste("Wildfire Smoke PM"[2.5]," 10µg/m"^3))) +
    # x label
    xlab("Date") +
    ggtitle(expression(paste('A. Wildfire Smoke PM'[2.5]))) +
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
                                        size = 0.5),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_text(angle = 90))

pm_msa_plot
```

Bind plots together.

```{r final_plot}
# save plot
#tiff('./r_scripts/analysis/Figs/msa_smoke_asthma.tiff', width = 600, height = 650)
gridExtra::grid.arrange(pm_msa_plot, pos_plot, ncol=2, widths = c(1,1))
#dev.off()
```

### Monthly Rate and Sum by MSA

```{r monthly_counts_msa}
monthly_msa_counts <- asthma_pos_count %>% 
  mutate(month = lubridate::month(date)) %>% 
  group_by(month, metroarea, service_place) %>% 
  summarise(n_events = sum(n_events)) %>% 
  left_join(population, by = 'metroarea') %>% 
  mutate(rate_100k = (n_events/pop)*100000,
         month_name = factor(case_when(month == 5 ~ 'May',
                                       month == 6 ~ 'Jun',
                                       month == 7 ~ 'Jul',
                                       month == 8 ~ 'Aug', 
                                       month == 9 ~ 'Sep'),
                        levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep'))) %>% 
  filter(service_place != 'Ambulance-Land') %>% 
  filter(service_place != 'Urgent Care Facility') %>% 
  filter(!is.na(pop)) %>% 
  arrange(service_place, metroarea, month)
```

Plot montly rate by MSA.

ED Rate

```{r ed_plot}
ed_count <- monthly_msa_counts %>% filter(service_place == 'Emergency Department')

# plot
ed_plot <- ggplot(ed_count, aes(x = month_name, group = metroarea,
                                        fill = metroarea)) +
  geom_bar(aes(weight = rate_100k)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5"),
                    guide = F) +
  facet_wrap(~metroarea, ncol =1) +
  ylab("Monthly Rate per 100,000 Persons") +
  xlab("Date") +
  ggtitle('A. ED') +
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
                                        size = 0.5),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_text(angle = 90),
        legend.key = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1))

ed_plot
```

Outpatient.

```{r outpat}
out_count <- monthly_msa_counts %>% 
  filter(service_place == 'Outpatient Hospital')
# plot
out_plot <- ggplot(out_count, aes(x = month_name, group = metroarea,
                                        fill = metroarea)) +
  geom_bar(aes(weight = rate_100k)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5"),
                    guide = F) +
  facet_wrap(~metroarea, ncol =1) +
  xlab("Date") +
  ggtitle('B. Outpatient') +
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
                                        size = 0.5),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_blank(),
        legend.key = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1))

out_plot
```

Inpatient.

```{r outpat}
inpat_count <- monthly_msa_counts %>% 
  filter(service_place == 'Inpatient Hospital')
# plot
inpat_plot <- ggplot(inpat_count, aes(x = month_name, group = metroarea,
                                        fill = metroarea)) +
  geom_bar(aes(weight = rate_100k)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5"),
                    guide = F) +
  facet_wrap(~metroarea, ncol =1) +
  ylab("Monthly Rate per 100,000 Persons") +
  xlab("Date") +
  ggtitle('C. Inpatient') +
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
                                        size = 0.5),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1))

inpat_plot
```

Office.

```{r office}
off_count <- monthly_msa_counts %>% 
  filter(service_place == 'Office')
# plot
off_plot <- ggplot(off_count, aes(x = month_name, group = metroarea,
                                        fill = metroarea)) +
  geom_bar(aes(weight = rate_100k)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5"),
                    guide = F) +
  facet_wrap(~metroarea, ncol =1) +
  ylab("Monthly Rate per 100,000 Persons") +
  xlab("Date") +
  ggtitle('D. Office') +
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
                                        size = 0.5),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_blank(),
        legend.key = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1))

off_plot
```

SABA fill.

```{r saba}
saba_count <- monthly_msa_counts %>% 
  filter(service_place == 'SABA Fill')
# plot
saba_plot <- ggplot(saba_count, aes(x = month_name, group = metroarea,
                                        fill = metroarea)) +
  geom_bar(aes(weight = rate_100k)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~metroarea, ncol =1) +
  ylab("Monthly Rate per 100,000 Persons") +
  xlab("Date") +
  ggtitle('E. SABA') +
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_line(color = 'grey', linetype = 'dotted',
                                        size = 0.5),
        panel.grid.minor = element_blank(),
        # strip element
        strip.background = element_rect(colour=NA, fill=NA),
        panel.border = element_rect(fill = NA, color = "black"),
        # facet text size
        strip.text = element_text(size = 10),
        axis.title.y = element_blank(),
        legend.key = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1))

saba_plot
```

Join all plots.

```{r msa_rate_plot}
# save plot
#tiff('./r_scripts/analysis/Figs/msa_month_rate.tiff', width = 800, height = 600)
gridExtra::grid.arrange(ed_plot, out_plot, inpat_plot, off_plot, saba_plot, 
                        ncol=5, widths = c(0.2, 0.2, 0.2, 0.2, 0.3))
#dev.off()
```

Compare August to July and September for each MSA.

Model
```{r rate_compare}
metro <- unique(monthly_msa_counts$metroarea)
service_place <- unique(monthly_msa_counts$service_place)
month <- seq(5,9,by=1)
# expand grid
month_msa <- expand.grid(metro, service_place, month) %>% 
  rename(metroarea = Var1, service_place = Var2, month = Var3)

summary(outcome_list)
# create outcome list
outcome_list <- month_msa %>% 
  left_join(monthly_msa_counts, by = c("metroarea", "service_place", "month")) %>% 
  mutate(n_events = ifelse(is.na(n_events), 0, n_events),
         pop = ifelse(is.na(pop), 85934, pop),
         month_name = factor(case_when(month == 5 ~ 'May',
            month == 6 ~ 'Jun',
            month == 7 ~ 'Jul',
            month == 8 ~ 'Aug', 
            month == 9 ~ 'Sep'),
            levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep'))) %>% 
  split(.$service_place)

# function list
compare_df <- outcome_list %>% 
  map_dfr(., function(df){
    df <- outcome_list[[5]]
    serv_pl <- unique(df$service_place)
    serv_pl
    # model
    mod <- glm(n_events ~ 0 + metroarea*month_name + offset(log(pop)), 
                family = poisson, data = df)

    summary(mod)
    contr_msa_jul_to_sep <- rbind(
      #'Bend: Aug to Jul' = c(0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
      #'Bend: Aug to Sept' = c(0,0,0,0,0,0,0,0,-1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
      #'Cor: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0,0,0),
      #'Cor: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0),
      #'Eug: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0,0),
      #'Eug: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0),
      'Med: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0,0,0),
      'Med: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0),
      'Port vs Med: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-.5,-.5,0,0,0,.5,.5,0,0,0,0,0,0)
      #'Port: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0),
      #'Sal: Aug to Jul' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1,0,0,0,0,0),
      #'Sal: Aug to Sept' = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,1)
      )

    # multicomp comparison of contrasts
    jul_sep_contrast <- multcomp::glht(mod, contr_msa_jul_to_sep)
    # summary of adjusted
    compare_df <- data.frame(broom::tidy(
      summary(jul_sep_contrast, test = multcomp::adjusted('none')))) %>% 
      mutate(service_place = serv_pl) %>% 
      select(service_place, lhs:p.value)
  })
compare_df
```

Test change by MSA.

```{r}
print(saba_count)
```

Fitting a time-series model to predict the observed across MSAs. I need the spline package.

```{r spline}
library(splines)
```

I want to map across each outcome and metroarea to fit splines I think. I'll first try an interaction spline by MSA for saba fills.

```{r weekly_pm}
weekly_pm = msa_pm %>% 
  mutate(week_date = lubridate::ceiling_date(date, unit = 'week')) %>% 
  group_by(metroarea, week_date) %>% 
  summarise(mu_gwr_smk = mean(geo_smk_pm), mu_gwr = mean(geo_wt_pm))

summary(weekly_pm)
```

```{r}
glimpse(asthma_weekly_count)

# vector of outcomes
out_vec = unique(asthma_weekly_count$service_place)

# list of metroareas
msa_list = asthma_weekly_count %>% 
  split(., list(.$metroarea,.$service_place))

# map to get estiamtes of counts for each 
exp_n = msa_list %>% 
  map_dfr(., function(data){
    service_place = as.character(data$service_place[1])
    metroarea = as.character(data$metroarea[1])
    # vector of dates
    week_date = as.character(data$week_date)
    # spline with 4 degrees of freedom fits msas pretty well
    spl = ns(data$week_date, df = 4)
    # fit spline to counts
    mod = glm(n_events ~  spl, family = 'poisson'(link='identity'), 
          data = data)
    # estimated outcomes
    est = predict(mod, newdata=data, type = 'response') 
    # dataframe of estimates by msa and place of service
    est_df = data.frame(cbind(service_place, metroarea, week_date, est)) %>% 
      mutate(week_date = as.Date(week_date, format = "%Y-%m-%d"),
             est = as.numeric(as.character(est)),
             metroarea = as.character(metroarea),
             service_place = as.character(service_place)) 
  }) %>% 
  left_join(asthma_weekly_count, ., 
            by = c('metroarea', 'service_place', 'week_date')) %>% 
  mutate(diff = est - n_events,
         est_rate = (est/pop)*100000,
         diff_rate = rate_100k - est_rate)

# plot of expected over observed
ggplot(data=exp_n, aes(x=week_date, y=diff_rate, 
    group = service_place, color = service_place)) +
  geom_point() +                 
  facet_wrap(~metroarea)
```

Saving dataframe to plot in python.

```{r save_df, eval=F}
# write asthma weekly count
#write_csv(asthma_weekly_count, './data/health/asthma_weekly_count.csv')
# write pm msa
#write_csv(msa_pm, './data/pm/msa_pm.csv')
#write_csv(monthly_msa_counts, './data/monthly_counts.csv')
```

Overlap plot.

```{r eval=F}
grid::grid.newpage()
vp1 <- grid::viewport(width = 1, height = 1, x = 0.5, y = 0.5)
vp2 <- grid::viewport(width = 1, height = 1, x = 0.5, y = 0.5)
print(pm_msa_plot, vp = vp1)
print(pos_plot, vp = vp2)
```

```{r}
exp(-0.295)
```