### 2018-09-17 r_scripts/analysis/wfs_exposure_figures.Rmd
Updated scripts for paper.

---
title: "Douglas-Complex Fire PM2.5"
author: "Ryan Gan"
date: "6/4/2018"
output: html_document
---

##1. Introduction

  In the summer of 2013, the Douglas-Complex fires occured in southwest Oregon. Oregonians in this part of the state were at risk of exposure to extreme levels of particulate matter (PM). This project aim is to determine if there is an association to smoke from the Douglas-Complex fires and acute cardiopulmonary morbidity in the state of Oregon.

Packages used: tidyverse, survival.
```{r setup, echo =F, warning=F, message=F}
library(tidyverse) # general data wrangle
library(survival) # conditional logistic regression
library(geofacet)

# knitr options
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
echo=FALSE, warning=FALSE, message=FALSE)
```

##2. Wildfire smoke descriptive characteristics

Smoke exposure is represented by ZIP code-level population-weighted PM~2.5~ estimated using geographically-weighted ridge regression (GWR), where we have subtracted off the background median PM~2.5~ for the month where no smoke was present overhead in the atmospheric column using HMS. It was estimated every day for the wildfire season of May to October, 2013.

We've decided to represent descriptive characteristics of smoke by county-level as we can represent the location and temporal variation of smoke exposure better than we could at the ZIP code level.

```{r smoke data import and wrangle}
# pm path
pm_path <- paste0("../../data/pm/2013-oregon_county_pm25.csv")
# county pm
county_pm_df <- read_csv(pm_path) %>% 
  mutate(county = stringr::str_replace(county, "[.]", ""))

# fire locations
fire_locations <- read_csv("../../data/pm/2013-oregon_firelocations.csv") %>% 
  mutate(area = sqrt(total_acres/1000))

# estimate county smoke 
county_smk_count <- county_pm_df %>% 
  # binary smoke variable at >10 ug/m^3
  mutate(gwr_smk10 = ifelse(geo_smk_pm >= 10, 1, 0),
         gwr_smk15 = ifelse(geo_smk_pm >= 15, 1, 0)) %>% 
  group_by(county) %>% 
  summarise(smk_day10 = sum(gwr_smk10), smk_day15 = sum(gwr_smk15)) %>% 
  # lower case county name and remove "." to match spatial df
  mutate(subregion = tolower(stringr::str_replace(county, "[.]", " ")))

# extract oregon map data
or_county_spatial_df <- map_data("county", "oregon") %>% 
  right_join(county_smk_count, by = "subregion")

# use the map function from the maps package to get the polygon data
county_poly <- maps::map("county", "oregon", plot=FALSE, fill = TRUE)
# find centroids
county_centroids <- maps:::apply.polygon(county_poly, maps:::centroid.polygon)

# create a data frame for graphing out of the centroids of each polygon
county_centroids <- county_centroids[!is.na(names(county_centroids))]
centroid_array <- Reduce(rbind, county_centroids)

county_text <- reduce(county_centroids, rbind) %>% 
  as_data_frame() %>% 
  rename(long = V1, lat = V2) %>% 
  mutate(county = stringr::str_sub(
    stringr::str_to_title(names(county_centroids)), start=8L))

# import MSA location
msa_path <- paste0("../../data/pm/msa_coordinate.csv")
msa_df <- read_csv(msa_path)

msa_location <- msa_df %>%
  filter(grepl('OR', msa_df$MSA)) %>%
  filter(MSA != "Albany, OR" &
         MSA != "Grants Pass, OR")

# create msa label data frame
msa_text <- msa_location %>%
  select(MSA, Longitude, Latitude)

msa_text$MSA <- gsub(",.*", "", msa_text$MSA)

msa_text <- msa_text %>% 
  mutate(metro_area = case_when(MSA == 'Portland-Vancouver-Hillsboro' ~ 'Portland',
                                MSA == 'Bend-Redmond' ~ 'Bend', 
                                T ~ MSA))
```

We summarize smoke exposure two ways: as a count of days where smoke PM~2.5~ is > 15 ug/m^3 (to increase the specificity to identify areas impacted by smoke) at the county level, and time-series of county-population weighted PM~2.5~ over May to October 2013.

### Number of days in Oregon where smoke PM~2.5~ exceeded 15 ug/m^3.

```{r smoke days map}
smoke_map <- ggplot(or_county_spatial_df, aes(x=long,y=lat, group=group)) +
  # fill with number of smoke days
  geom_polygon(aes(fill = smk_day15)) +
  scale_fill_gradient2(expression("Smoke Days > 15 µg/m"^3),
                       low = "white", mid= "#00dbde", high="#6a3093", 
                       midpoint = 10,
    guide = guide_colorbar(direction = "horizontal", title.position = "top",
                          title.vjust = 2, title.hjust = 0.5, barwidth = 15)) +
  # add county path on top
  geom_path(colour="#4ac29a") +
  # adding fire locations in
  geom_point(data=fire_locations, 
             aes(x=lon, y=lat, size = sqrt(area/2)), 
             color = "red", shape = 17, alpha = 0.7, 
             show.legend = T, inherit.aes = FALSE) +
  # legend for firsts for fires
  scale_size('Wildfire Acres Burned', breaks = c(0.8, 1, 1.6), 
             labels = c(1600, 4000, 26000),
             guide = guide_legend(title.position = "top", direction = "horizontal",
                                  title.hjust = 0.5, barwidth = 10)) +
  # add county text 
  geom_text(data = county_text, 
            aes(x=long, y=lat, label = county, group="county"), 
            color = "#33001b", size = 2.5, inherit.aes = F) +
  # add metropolitan
  geom_point(data = msa_location, 
    aes(x=Longitude, y=Latitude), 
    size = 1, color = '#ff0084', shape = 19, alpha = 0.7, 
    show.legend = F, inherit.aes = FALSE) +
  # add metro text
  geom_text(data = msa_text, 
    aes(x=Longitude + c(0.15,0.1,0.1,-0.05,0.2,0.1), 
        y=Latitude + c(0.2,0.1,0.1,-0.1,0.1,0.1), 
        label = metro_area, group="MSA"), 
    color = "#ff0084", size = 3) +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(colour=NA, fill=NA),
        legend.position = "bottom",
        legend.key = element_rect(fill = NA, colour = NA, size = 0.25))

smoke_map

# save figure
 ggsave(filename = "./r_scripts/analysis/Figs/smoke_map.tiff", plot = smoke_map,
        units="in", width=7, height=6, dpi=300)
```

Most of the heavy smoke days were in the south-west part of Oregon. Triangles are fires; had them in the legend, but the legend looked weird so I took them out.

### Time-series of geo-weighted regression estimates of smoke PM~2.5~ by county

Using the geofacet package to organize small-multiples by geography.

*Note 11/7/17: Not sure how much I love the layout of this Oregon geofacet grid, but we'll see what people think. I like that geofacet lets you look at the two maps together to understand the exposure pattern, but too much whitespace. However, with this orientation, I believe you could cut out the smoke days figure.*

```{r geofacet time series of pm}
# use or_counties_grid from geo_facet 
or_grid <- geofacet::us_or_counties_grid1 %>% 
mutate(name = ifelse(name == "HoodRiver", "Hood River", name))

# small multiples plot
plot <- ggplot(county_pm_df, aes(x=date, y= geo_smk_pm)) + 
  geom_point(color = "#6a3093", size = 0.5) +
  #scale_x_date(date_breaks = "1 month", labels = "%b") +
  geofacet::facet_geo(~county, grid = or_grid) +
  ylab(expression(paste("Wildfire Smoke PM"[2.5], " µg/m"^3))) +
  xlab("2013 Wildfire Season") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    strip.text = element_text(size = 8),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(colour=NA, fill=NA),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0))
# plot 
plot
# save figure
# ggsave(filename = "./r_scripts/analysis/Figs/county_smoke_timeseries.tiff", 
#        plot = plot, units="in", width=6, height=5, dpi=300)
```